---
title: "Adaptive potential\nModels"
author: "Bioshifts group"
output:   
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

# Load libraries and source code

```{r libs, message=FALSE}

rm(list=ls())

list.of.packages <- c("ggplot2", "data.table", "dplyr", "tidyr", "parallel", "bdc", "taxadb", "traitdataform", "pbapply", "tidyverse", "readxl", "lme4", "coefplot", "sjPlot", "sjmisc", "effects", "rgdal", "maptools", "terra", "MuMIn", "lsmeans", "GGally", "tidyterra", "glmmTMB", "DHARMa") 

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]

if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)
```

# Load data

```{r splist}

# Use only Fonseca
# mydataset <- read.csv(here::here("Data/gen_data_final_m5.csv"))
mydataset <- read.csv(here::here("Data/gen_data_v1_dist2_fon.csv"))

# how many data at each shift gradient
mydataset %>%
  group_by(Type) %>%
  tally

# how many data at each shift gradient / parameter
mydataset %>%
  group_by(Type,Param) %>%
  tally

summary(mydataset$velocity[which(mydataset$shift_vel_sign=="pospos")])
summary(mydataset$velocity[which(mydataset$shift_vel_sign=="negneg")])

summary(mydataset$SHIFT[which(mydataset$shift_vel_sign=="pospos")])
summary(mydataset$SHIFT[which(mydataset$shift_vel_sign=="negneg")])

summary(mydataset$velocity[which(mydataset$SHIFT>0)])
summary(mydataset$velocity[which(mydataset$SHIFT<0)])

```

# Data description

**Genetic data**  
- He = heterozigosite measurements from multiple data sources (DeKort et al 2021, Nature Communications; Lawrence and Fraser 2020, Global Ecology and Biogeography; Malin, Unpublished).  
Genetic data was harmonized following multiple methods, so that measurements across different marker types are comparable:  
- He_harm: transformation used in DeKort et al 2021 (Nature Communications). Makes markers range from 0-1.
1) scale values at each marker type  
2) standardize (x - min)/(max-min)  
- He_harm2: transformation suggested by Malin. Applies the function transform01 before a logit transformation. 
1) transform01 >> The only intent of this function is to move slightly inward the values of He so that it allows applying logit transformation.
2) apply logit transformation  
- He_harm3: He_harm2 centered in zero.   

Genetic data was paired with bioshifts data based on geographical distance of observations. For each species range shift, get the closest He to the centroid (for O), to the max latitude (for LE, at the North hemisphere), and to the min latitude (for TE, at the North hemisphere) of the study area.   

We did the same procedure but averaging He values and weighting by distance from the range shift observation. We flag He_harm metrics generated with the average distance procedure with the suffix "-w".  

**Shift data**  
We ran models using raw and *corrected* shift values. Corrected shifts were estimated based on models that account for variability in methodology of the different studies estimating range shifts and taxonomy.  
- SHIFT = raw data  
- SHIFT_cor = corrected data  

**Lag data**  
We calculated lag as the difference between the velocity of temperature change and that of species range shift (lag = velocity - shift). A positive value means that a species range shift lag behind climate change velocity, values close to zero means species track climate change closely, whereas a negative value means species are shift their range faster than expected by the velocity of climate change (overshoot). In cases the velocity of climate change was negative, we change the sign of lag, so that positive values always mean lag, and negative values always mean overshoot.  
We tried to account and explore for situations that may generate confusion, as described the metrics bellow:  
- lag: velocity - shift (we change the sign of lag when velocity is negative)  
- lag2: when species shift in the opposite direction of velocity, lag = velocity  
- lag3: when shift is faster than velocity, lag = 0.  


# Select only shifts in the same direction of velocity
```{r}

mydataset <- mydataset %>%
  filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")
# 
# summary(mydataset$SHIFT[which(mydataset$shift_vel_sign=="pospos")])
# summary(mydataset$SHIFT[which(mydataset$shift_vel_sign=="negneg")])
# 
# summary(mydataset$velocity[which(mydataset$shift_vel_sign=="pospos")])
# summary(mydataset$velocity[which(mydataset$shift_vel_sign=="negneg")])


```

# Select groups with at least 5 observations
```{r}

n_groups = 5

type_groups_lat <- mydataset %>%
  filter(Type == "LAT") %>%
  group_by(Group,Param) %>%
  dplyr::summarise(N = length(Param)) %>%
  dplyr::filter(N>n_groups) %>%
  group_by(Group) %>%
  dplyr::summarise(N = length(Param)) %>%
  dplyr::filter(N>=3)

type_groups_ele <- mydataset %>%
  filter(Type == "ELE") %>%
  group_by(Group,Param) %>%
  dplyr::summarise(N = length(Param)) %>%
  dplyr::filter(N>n_groups) %>%
  group_by(Group) %>%
  dplyr::summarise(N = length(Param)) %>%
  dplyr::filter(N>=3)

mydataset <- mydataset %>%
  dplyr::filter((Type == "LAT" & Group %in% type_groups_lat$Group) | (Type == "ELE" & Group %in% type_groups_ele$Group))

```

# Remove outliers in shift

```{r message=FALSE}

# LAT
LAT_obs <- mydataset %>% 
  filter(Type == "LAT")

summary(LAT_obs$SHIFT_abs)
hist(LAT_obs$SHIFT_abs)

thr <- quantile(LAT_obs$SHIFT_abs,c(.01,0.99))
thr

rem <- which((LAT_obs$SHIFT_abs > thr[1]) & (LAT_obs$SHIFT_abs < thr[2]))
if(any(rem)){
  LAT_obs <- LAT_obs[rem,]
}
summary(LAT_obs$SHIFT_abs)
hist(LAT_obs$SHIFT_abs)

hist(sqrt(LAT_obs$SHIFT_abs))


# ELE
ELE_obs <- mydataset %>% 
  filter(Type == "ELE")

summary(ELE_obs$SHIFT_abs)
hist(ELE_obs$SHIFT_abs)

thr <- quantile(ELE_obs$SHIFT_abs,c(.01,0.99))
thr

rem <- which((ELE_obs$SHIFT_abs > thr[1]) & (ELE_obs$SHIFT_abs < thr[2]))
if(any(rem)){
  ELE_obs <- ELE_obs[rem,]
}
summary(ELE_obs$SHIFT_abs)
hist(ELE_obs$SHIFT_abs)

hist(sqrt(ELE_obs$SHIFT_abs))

# Combine
mydataset <- rbind(LAT_obs,ELE_obs)
hist(mydataset$SHIFT_abs)
hist(sqrt(mydataset$SHIFT_abs))

##############
# Lags look ok

# remove outliers in lag
# summary(mydataset$lag)
hist(mydataset$lag)
# thr <- quantile(mydataset$lag,c(.01,0.99))
# thr
# 
# rem <- which((mydataset$lag > thr[1]) & (mydataset$lag < thr[2]))
# if(any(rem)){
#     mydataset <- mydataset[rem,]
# }
# summary(mydataset$lag)
# hist(mydataset$lag)
```


# Playing with models

## Latitude

```{r message=FALSE}

mydatatogo <- mydataset  %>%
  dplyr::filter(Type == "LAT") %>%
  mutate(
    # He_harm = scale(He_harm),
    # He_harm2 = scale(He_harm2),
    # He_harm3 = scale(He_harm3),
    # He_harm_w = scale(He_harm),
    # He_harm3_w = scale(He_harm3_w),
    # He_harm3_w = scale(He_harm3_w),
    velocity_raw = velocity,
    velocity = scale(velocity_raw),
    velocity_abs = scale(abs(velocity_raw)),
    lag = scale(lag),
    lag2 = scale(lag2),
    lagC = scale(lagC),
    lagC2 = scale(lagC2),
    lagC3 = lagC3,
    ID.area = scale(ID.area),
    DUR = scale(DUR),
    dist_weight = 1/log(min_dist), # this is the transformation I found leads to a normal distribution
    START = scale(START),
    SHIFT_raw = SHIFT,
    SHIFT = scale(SHIFT_raw),
    SHIFT_abs = 
      # scale(sqrt(
      SHIFT_abs
    # ))
    ,
    SHIFT_cor_raw = SHIFT_cor,
    SHIFT_cor = scale(SHIFT_cor_raw),
    SHIFT_cor_abs = abs(SHIFT_cor),
    Param = factor(Param),
    Group = factor(Group),
    spp = factor(spp)) %>%
  dplyr::select(
    He,
    # He_harm, He_harm3, He_harm3, 
    # He_harm_w, He_harm3_w, He_harm3_w, 
    lag, lag2, lag3, lagC, lagC2, lagC3, ECO,
    SHIFT, SHIFT_abs, SHIFT_cor, SHIFT_cor_abs, SHIFT_raw, SHIFT_cor_raw,
    Param, Group, spp, Class, Family, Genus, velocity, velocity_abs, velocity_raw,
    START, DUR, Uncertainty_Parameter, Uncertainty_Distribution, Grain_size, Data, ID.area, Sampling, Article_ID,
    min_dist, dist_weight,
    Marker,
    shift_vel_sign)

mydatatogo <- data.frame(mydatatogo)
dim(mydatatogo)

hist(mydatatogo$SHIFT_cor_abs)
hist(mydatatogo$SHIFT_cor_raw)

ggplot(mydatatogo %>%
         filter(ECO=="T"),
       aes(x=He,y=SHIFT,color = velocity, size = velocity))+
  geom_point()+
  theme_bw()

ggplot(mydatatogo %>%
         filter(ECO=="M"),
       aes(x=He,y=SHIFT,color = velocity, size = velocity))+
  geom_point()+
  theme_bw()

```


### Shift model
```{r fig.height=10, fig.width=10}

model_data = mydatatogo

# define the He metric to go
# model_data$He <- as.numeric(model_data$He)
# model_data$SHIFT <- as.numeric(model_data$SHIFT)
# model_data$SHIFT <- as.numeric(model_data$SHIFT_raw)

table(model_data$Group)

# All
bestmodel <- lmer(SHIFT_abs ~ 
                    # Fixed effects
                    ## Ecology
                    velocity_abs*He + Param +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    (velocity_abs|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data,
                  na.action = "na.omit")

# Test with methods in the random intercept
# Use genus as random intercept


# All
# levels(model_data$Param)
model_data$Param <- relevel(model_data$Param, ref = "O") 

bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He*Param +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    (velocity_abs|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data,
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Fish"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

# plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity_abs:He`)
plot(tmp$`velocity:He:Param`)

summary(bestmodel)
confint(bestmodel, method = "Wald")

# Fish
bestmodel <- lmer(SHIFT_abs ~ 
                    # Fixed effects
                    ## Ecology
                    velocity_abs*He +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Fish"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Fish"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

# plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity_abs:He`)

summary(bestmodel)
confint(bestmodel)


# Insects
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He +
                    ## Control Methods
                    # DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Insect"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    # DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Insect"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

# plot(tmp$`velocity:He:Param`)
plot(tmp$`velocity:He`)

summary(bestmodel)
confint(bestmodel)

## Plants
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He*Param +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (velocity_abs|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data,
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Plant"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity:He:Param`)

summary(bestmodel)
confint(bestmodel)



## Birds
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Bird"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Bird"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity:He`)

summary(bestmodel)
confint(bestmodel)



```

## Elevation

```{r message=FALSE}

mydatatogo <- mydataset  %>%
  dplyr::filter(Type == "ELE") %>%
  mutate(
    # He_harm = scale(He_harm),
    # He_harm2 = scale(He_harm2),
    # He_harm3 = scale(He_harm3),
    # He_harm_w = scale(He_harm),
    # He_harm3_w = scale(He_harm3_w),
    # He_harm3_w = scale(He_harm3_w),
    velocity_raw = velocity,
    velocity = scale(velocity_raw),
    velocity_abs = scale(abs(velocity_raw)),
    lag = scale(lag),
    lag2 = scale(lag2),
    lagC = scale(lagC),
    lagC2 = scale(lagC2),
    lagC3 = lagC3,
    ID.area = scale(ID.area),
    DUR = scale(DUR),
    dist_weight = 1/log(min_dist), # this is the transformation I found leads to a normal distribution
    START = scale(START),
    SHIFT_raw = SHIFT,
    SHIFT = scale(SHIFT_raw),
    SHIFT_abs = 
      # scale(sqrt(
      SHIFT_abs
    # ))
    ,
    SHIFT_cor_raw = SHIFT_cor,
    SHIFT_cor = scale(SHIFT_cor_raw),
    SHIFT_cor_abs = abs(SHIFT_cor),
    Param = factor(Param),
    Group = factor(Group),
    spp = factor(spp)) %>%
  dplyr::select(
    He,
    # He_harm, He_harm3, He_harm3, 
    # He_harm_w, He_harm3_w, He_harm3_w, 
    lag, lag2, lag3, lagC, lagC2, lagC3, ECO,
    SHIFT, SHIFT_abs, SHIFT_cor, SHIFT_cor_abs, SHIFT_raw, SHIFT_cor_raw,
    Param, Group, spp, Class, Family, Genus, velocity, velocity_abs, velocity_raw,
    START, DUR, Uncertainty_Parameter, Uncertainty_Distribution, Grain_size, Data, ID.area, Sampling, Article_ID,
    min_dist, dist_weight,
    Marker,
    shift_vel_sign)

mydatatogo <- data.frame(mydatatogo)
dim(mydatatogo)

hist(mydatatogo$SHIFT_cor_abs)
hist(mydatatogo$SHIFT_cor_raw)

ggplot(mydatatogo %>%
         filter(ECO=="T"),
       aes(x=He,y=SHIFT,color = velocity, size = velocity))+
  geom_point()+
  theme_bw()


```


### Shift model Param
```{r fig.height=10, fig.width=10}

model_data = mydatatogo

# define the He metric to go
# model_data$He <- as.numeric(model_data$He)
# model_data$SHIFT <- as.numeric(model_data$SHIFT)
# model_data$SHIFT <- as.numeric(model_data$SHIFT_raw)

table(model_data$Group)



# Insects
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Insect"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    # DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Insect"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

# plot(tmp$`velocity:He:Param`)
plot(tmp$`velocity:He`)

summary(bestmodel)
confint(bestmodel)

## Plants
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He*Param +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Plant"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Plant"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity:He`)

summary(bestmodel)
confint(bestmodel)



## Birds
bestmodel <- lmer(SHIFT ~ 
                    # Fixed effects
                    ## Ecology
                    velocity*He*Param +
                    ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Bird"),
                  na.action = "na.omit")

baseline <- lmer(SHIFT ~ 
                   # Fixed effects
                   ## Ecology
                   velocity +
                   ## Control Methods
                    DUR + ID.area + Grain_size + Data +
                    # Random effects
                    ## Taxonomy
                    # (1|Group) +
                    ## Studies
                    (1|Article_ID),
                  # weight by distance of genetic data to range shift data
                  weights = dist_weight,
                  data = model_data %>%
                    filter(Group=="Bird"),
                  na.action = "na.omit")


AIC(bestmodel,baseline)

tmp <- effects::allEffects(bestmodel)

plot(tmp$`He:velocity:Param`)
plot(tmp$`velocity:He`)

summary(bestmodel)
confint(bestmodel)



```

