---
title: "Adaptive potential\nMerge genetic data"
author: "Bioshifts group"
output:   
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

# Load libraries and source code

```{r libs, message=FALSE}
rm(list=ls())

library(ggplot2)
require(data.table)
library(dplyr)
library(tidyr)
require(parallel)
library(bdc)
library(taxadb)
library(traitdataform)
library(pbapply)
library(tidyverse)
library(readxl)
library(lme4)
library(coefplot)
library(sjPlot)
library(sjmisc)
library(effects)
library(rgdal)
library(pbapply)
library(maptools)
library(rgeos)
library(terra)
library(nlme)
library(pscl)
library(lsmeans)
library(arm)
library(MCMCglmm)
library(PerformanceAnalytics)
library(GGally)
library(broom.mixed)
library(tidyterra)
library(rnaturalearth)

```

# Functions
```{r func}

# find accepted species names
source(here::here("R/clean_taxa_functions.r"))

# Malin's transformation for moving values slightly inward. 
transform01 <- function(x) (x * (length(x) - 1) + 0.5) / (length(x))
# De Kort transformation
deKort_trans <- function(p){
    p <- scale(p)
    p <- (p - min(p, na.rm = T))/(max(p, na.rm = T)-min(p, na.rm = T))
    return(p)
}
# Malin's transformation before a logit transformation
logit_trans <- function(p){
    p <- transform01(p)
    p <- log(p/(1-p))
    return(p)
}

# subset data for modelling
# Type: "LAT" or "ELE"
# n_groups: min number of species in a group to keep. If a group has < then n_groups, that groups is removed from the dataset. We do not want to fit models for a group with just a few species (e.g., 3 bird species) because species is used as a random effect.
get_data_to_go <- function(mydataset, Type, n_groups = 3){
    mydatatogo <- mydataset %>%
        dplyr::filter(Type == Type)
    
    type_groups <- mydatatogo %>%
        group_by(Group,Param) %>%
        dplyr::summarise(N = length(Param)) %>%
        dplyr::filter(N>n_groups) %>%
        group_by(Group) %>% 
        dplyr::summarise(N = length(Param)) %>%
        dplyr::filter(N>=n_groups)
    
    mydatatogo <- mydatatogo %>%
        dplyr::filter(Group %in% type_groups$Group) %>%
        mutate(He_harm = scale(He_harm),
               He_harm3 = scale(He_harm3),
               velocity = scale(velocity),
               lag = scale(lag),
               ID.area = scale(ID.area),
               DUR = scale(DUR),
               dist_weight = 1/log(min_dist), # this is the transformation I found leads to a normal distribution
               START = scale(START),
               SHIFT = scale(SHIFT),
               Param = factor(Param),
               Group = factor(Group),
               spp = factor(spp),
               Method = factor(paste(Uncertainty_Parameter, Uncertainty_Distribution, Grain_size, Data))) %>%
        dplyr::select(He_harm, He_harm3, lag, SHIFT, Param, Group, velocity, spp, START, DUR, Method, ID.area, Sampling, min_dist, dist_weight, Marker) %>%
        na.omit()
    
    return(mydatatogo)
}

```

# Load species list

```{r splist}

splist <- read.csv(here::here("Data/splist.csv"), header = T)
# remove duplicated sp_names
splist <- splist %>%
    dplyr::select(scientificName,kingdom,phylum,class,order,family,db) %>%
    dplyr::filter(!duplicated(scientificName))

```

# Load bioshifts

## Load v1

```{r v1}
biov1 <- read.csv(here::here("Data/biov1_fixednames.csv"), header = T)
# Fix references in biov1
biov1$Article_ID <- gsub("[^0-9.-]", "", biov1$Article_ID)
biov1$Article_ID <- as.numeric(biov1$Article_ID)
biov1$sp_name_std_v1 <- gsub("_"," ",biov1$sp_name_std_v1)
biov1 <- biov1 %>%
    dplyr::select(ID,Article_ID,Study_ID,Type,Param,Trend,SHIFT,UNIT,DUR,v.lat.mean,v.ele.mean,START,DUR,
                  START,END,group,ECO,Sampling,Uncertainty_Distribution,Uncertainty_Parameter,
                  Grain_size,Data,ID.area,
                  Phylum,Class,Order,Family,Genus,sp_name_std_v1) %>% # select columns
    mutate(
        Type = case_when(
            Type=="HOR" ~ "LAT",
            TRUE ~ as.character(Type)),
        Data = case_when(
            Data=="occurence-based" ~ "occurrence-based",
            TRUE ~ as.character(Data)),
        Species = sp_name_std_v1) %>% 
    mutate( # from m/year to km/year
        SHIFT = case_when(
            UNIT == "m/year" ~ SHIFT/1000, 
            UNIT == "km/year" ~ SHIFT, 
            TRUE ~ NA_real_))

biov1 <- biov1 %>%
    dplyr::filter(!is.na(sp_name_std_v1))

biov1$spp <- biov1$sp_name_std_v1

all(biov1$sp_name_std_v1 %in% splist$scientificName)

```

## Classify species

```{r classify}

# Class species as Terrestrial, Marine or Freshwater
Terv1 <- unique(biov1$sp_name_std_v1[which(biov1$ECO == "T")])
Terrestrials = unique(c(Terv1))

Mar1 <- unique(biov1$sp_name_std_v1[which(biov1$ECO == "M")])
Marine = unique(c(Mar1))

# Freshwater fish
FFishv1 <- unique(biov1$sp_name_std_v1[(biov1$Class == "Actinopterygii" | biov1$Class == "Cephalaspidomorphi") & biov1$ECO == "T"])
FFish = unique(c(FFishv1))
# Marine fish
MFishv1 <- unique(biov1$sp_name_std_v1[biov1$Class == "Actinopterygii" & biov1$ECO == "M"])
MFish = unique(c(MFishv1))

splist$ECO = NA
splist$ECO[which(splist$scientificName %in% Terrestrials)] <- "T"
splist$ECO[which(splist$scientificName %in% Marine)] <- "M"
splist$ECO[which(splist$scientificName %in% MFish)] <- "M"

biov1$ECO = NA
biov1$ECO[which(biov1$sp_name_std_v1 %in% Terrestrials)] <- "T"
biov1$ECO[which(biov1$sp_name_std_v1 %in% Marine)] <- "M"
biov1$ECO[which(biov1$sp_name_std_v1 %in% MFish)] <- "M"

splist$Group = NA
splist$Group[which(splist$class == "Phaeophyceae")] <- "Chromista"
splist$kingdom[which(splist$class == "Phaeophyceae")] <- "Chromista"
splist$Group[which(splist$phylum == "Rhodophyta")] <- "Seaweed"
splist$kingdom[which(splist$phylum == "Rhodophyta")] <- "Plantae"
splist$Group[which(splist$family == "Elminiidae")] <- "Barnacles"
splist$Group[which(splist$kingdom == "Bacteria")] <- "Bacteria"
splist$Group[which(splist$class == "Holothuroidea")] <- "Sea cucumber"
splist$Group[which(splist$class == "Aves")] <- "Bird"
splist$Group[which(splist$class == "Insecta")] <- "Insect"
splist$Group[which(splist$class == "Mammalia")] <- "Mammal"
splist$Group[which(splist$class == "Arachnida")] <- "Spider"
splist$kingdom[which(splist$kingdom == "Viridiplantae")] <- "Plantae"
splist$kingdom[which(splist$phylum == "Tracheophyta")] <- "Plantae"
splist$Group[which(splist$kingdom == "Plantae")] <- "Plant"
splist$Group[which(splist$class == "Hydrozoa")] <- "Hydrozoa"
splist$Group[which(splist$class == "Anthozoa")] <- "Sea anemones and corals"
splist$Group[which(splist$class == "Polychaeta")] <- "Polychaetes"
splist$Group[which(splist$phylum == "Mollusca")] <- "Molluscs"
splist$Group[which(splist$class == "Malacostraca")] <- "Crustacean"
splist$Group[which(splist$class == "Hexanauplia")] <- "Crustacean"
splist$Group[which(splist$class == "Maxillopoda")] <- "Crustacean"
splist$Group[which(splist$class == "Ostracoda")] <- "Crustacean"
splist$Group[which(splist$class == "Branchiopoda")] <- "Crustacean"
splist$Group[which(splist$class == "Asteroidea")] <- "Starfish"
splist$Group[which(splist$class == "Ascidiacea")] <- "Ascidians tunicates and sea squirts"
splist$class[which(splist$class == "Actinopteri")] <- "Actinopterygii"
splist$Group[which(splist$class == "Actinopterygii")] <- "Fish"
splist$Group[which(splist$class == "Elasmobranchii")] <- "Fish"
splist$Group[which(splist$order == "Perciformes")] <- "Fish"
splist$Group[which(splist$class == "Chondrichthyes")] <- "Fish"
splist$Group[which(splist$class == "Holocephali")] <- "Fish"
splist$Group[which(splist$class == "Cephalaspidomorphi")] <- "Fish"
splist$Group[which(splist$class == "Echinoidea")] <- "Sea urchin"
splist$Group[which(splist$class == "Crinoidea")] <- "Crinoid"
splist$Group[which(splist$class == "Holothuroidea")] <- "Sea cucumber"
splist$Group[which(splist$class == "Reptilia")] <- "Reptile"
splist$Group[which(splist$order == "Squamata")] <- "Reptile"
splist$Group[which(splist$class == "Ophiuroidea")] <- "Brittle stars"
splist$Group[which(splist$class == "Chilopoda")] <- "Centipedes"
splist$Group[which(splist$class == "Diplopoda")] <- "Millipedes"
splist$Group[which(splist$class == "Amphibia")] <- "Amphibian"
splist$Group[which(splist$kingdom == "Fungi")] <- "Fungi"
splist$Group[which(splist$order == "Balanomorpha")] <- "Barnacles"
splist$Group[which(splist$phylum == "Nematoda")] <- "Nematodes"
splist$Group[which(splist$class == "Myxini")] <- "Hagfish"
splist$Group[which(splist$kingdom == "Chromista")] <- "Chromista"

######################################
biov1 <- merge(biov1[,-which(names(biov1) %in% c("Group","ECO"))], 
               splist[,c("Group","ECO","scientificName")], 
               by.x = "sp_name_std_v1", by.y = "scientificName", 
               all.x = T)

```

## Filter LAT / ELE shifts
Use only latitudinal and elevational shifts
```{r filter}

# v1
biov1 <- biov1 %>%
    dplyr::filter(Type %in% c("ELE","LAT"))  # Use LAT ELE shifts

# splist
sps <- unique(biov1$sp_name)
splist <- splist %>% dplyr::filter(scientificName %in% sps)

```

## Remove species identified to the genus level or cf.

```{r clean}

if(any(grep("sp[.]",biov1$sp_reported_name_v1))){
    biov1 <- biov1 %>% dplyr::filter(!grepl("sp[.]",sp_reported_name_v1))
}
if(any(grep("sp[.]",biov1$sp_name_std_v1))){
    biov1 <- biov1 %>% dplyr::filter(!grepl("sp[.]",sp_name_std_v1))
}
if(any(grep("cf[.]",biov1$sp_reported_name_v1))){
    biov1 <- biov1 %>% dplyr::filter(!grepl("cf[.]",sp_reported_name_v1))
}
if(any(grep("cf[.]",biov1$sp_name_std_v1))){
    biov1 <- biov1 %>% dplyr::filter(!grepl("cf[.]",sp_name_std_v1))
}

#remove freshwater fishes
biov1 <- biov1[-which((biov1$Class == "Actinopterygii" | biov1$Group == "Fish") & biov1$ECO=="T"),]

#remove marine birds
biov1 <- biov1[-which(biov1$Class == "Aves" & biov1$ECO=="M"),]


if(any(grep("sp[.]",splist$scientificName))){
    splist <- splist %>% dplyr::filter(!grepl("sp[.]",scientificName))
}
if(any(grep("sp[.]",splist$scientificName))){
    splist <- splist %>% dplyr::filter(!grepl("sp[.]",scientificName))
}
if(any(grep("cf[.]",splist$scientificName))){
    splist <- splist %>% dplyr::filter(!grepl("cf[.]",scientificName))
}
if(any(grep("cf[.]",splist$scientificName))){
    splist <- splist %>% dplyr::filter(!grepl("cf[.]",scientificName))
}


```

## Reclassify methodological variables
```{r reclass_meth}

unique(biov1$Data)
biov1$Data[biov1$Data=="occurence-based"] <- "occurrence-based"
biov1$Data <- factor(biov1$Data, levels = unique(biov1$Data))
table(biov1$Data)

unique(biov1$Sampling)
biov1$Sampling = ifelse(biov1$Sampling %in% c("IRR","MULTIPLE(continuous)"),"MULTIPLE", biov1$Sampling)
biov1$Sampling <- ordered(biov1$Sampling, 
                          levels = c("TWO","MULTIPLE","CONTINUOUS"))
table(biov1$Sampling)

unique(biov1$Grain_size)
biov1$Grain_size <- ifelse(biov1$Grain_size %in% c("large","very_large"),"large",biov1$Grain_size)
biov1$Grain_size <- ordered(biov1$Grain_size, 
                            levels = c("small","moderate","large"))
table(biov1$Grain_size)

unique(biov1$Uncertainty_Distribution)
biov1$Uncertainty_Distribution <- ifelse(biov1$Uncertainty_Distribution %in% c("RESAMPLING","RESAMPLING(same)"),"RESAMPLING",
                                         ifelse(biov1$Uncertainty_Distribution %in% c("MODEL","MODEL+RESAMPLING(same)","RESAMPLING+MODEL"),"MODEL",
                                                ifelse(biov1$Uncertainty_Distribution %in% c("DETECTABILITY","RESAMPLING(same)+DETECTABILITY"),"DETECTABILITY",
                                                       biov1$Uncertainty_Distribution)))
biov1$Uncertainty_Distribution <- ifelse(biov1$Uncertainty_Distribution == "RAW","OPPORTUNISTIC","PROCESSED")  
table(biov1$Uncertainty_Distribution)

#transform study area
biov1$ID.area <- log(biov1$ID.area)

```

## Direction of shift
Identify direction of shift
```{r shift_direction}

biov1 <- biov1 %>%
    mutate(shift_sign = ifelse(SHIFT>0,"pos","neg"),
           velocity = ifelse(Type == "LAT", v.lat.mean, v.ele.mean),
           vel_sign = ifelse(velocity>0,"pos","neg"),
           shift_vel_sign = paste0(shift_sign,vel_sign))

ggplot(biov1, aes(x=shift_vel_sign))+
    geom_bar()+
    coord_flip()+
    facet_wrap(Type~Param)

```

## Get centroids of study areas
This will be used to merge genetic data with bioshifts based on distance of observations
```{r get_centroids}

# The input file geodatabase
# fgdb <- "C:/Users/brunn/NextCloud/Bioshifts/Study_Areas.gdb"
# 
# # List all feature classes in a file geodatabase
# fc_list <- ogrListLayers(fgdb)
# 
# # Get centroids
# cl <- makeCluster(detectCores()-1)
# clusterExport(cl, c("readOGR", "gCentroid", "fgdb"))
# 
# centroids <- pblapply(fc_list, function(x){
#     fc <- readOGR(dsn=fgdb,layer=x)
#     c <- data.frame(gCentroid(fc))
#     names(c) <- c("centroid.x", "centroid.y")
#     
#     geomet <- data.frame(terra::geom(terra::vect(fc)))
#     
#     max_lat <- order(geomet[,"y"], decreasing = T)[1]
#     max_lat <- geomet[max_lat,c("x","y")]
#     names(max_lat) <- c("max_lat.x", "max_lat.y")
#     
#     min_lat <- order(geomet[,"y"])[1]
#     min_lat <- data.frame(geomet[min_lat,c("x","y")])
#     names(min_lat) <- c("min_lat.x", "min_lat.y")
#     
#     cbind(c, max_lat, min_lat)
# }, cl = cl)
# 
# stopCluster(cl)
# 
# centroids <- do.call("rbind",centroids)
# centroids$ID <- fc_list
# 
# write.csv(centroids, "Data/centroids_study_areas.csv", row.names = FALSE)

centroids <- read.csv(here::here("Data/centroids_study_areas.csv"))

biov1 <- merge(biov1, centroids, by = "ID")

```

# Calculate lags 

```{r lags}

# calculate lags
# positive values are a lag (range shift lower smaller than expected or in opposite directions) and negative values are range shift larger than expected
biov1$lag <- biov1$velocity-biov1$SHIFT
position = which(biov1$vel_sign == "neg")
biov1$lag[position] <- biov1$lag[position] * -1 # Any negative velocity means the sign of lag has to shift.

# Lag 2
# calculate lags based on "corrected" shifts
# biov1$lag2 <- biov1$velocity-biov1$shift_corrected
# position = which(biov1$vel_sign == "neg")
# biov1$lag2[position] <- biov1$lag2[position] * -1 # Any negative velocity means the sign of lag has to shift.

# Lag 3
# When shift is in opposite sign of velocity, lag == velocity
biov1$lag3 = biov1$lag
position <- which(biov1$shift_vel_sign == "posneg" | biov1$shift_vel_sign == "negpos")
biov1$lag3[position] <- biov1$velocity[position]

# Lag 4
# apply lag 3 to lag 2
# biov1$lag4 = biov1$lag2
# position <- which(biov1$shift_vel_sign == "posneg" | biov1$shift_vel_sign == "negpos")
# biov1$lag4[position] <- biov1$velocity[position]


# {
# par(mfrow=c(2,2))
# plot(lag~velocity, biov1)
# plot(lag2~velocity, biov1)
# plot(lag3~velocity, biov1)
# plot(lag4~velocity, biov1)
# }
# 
# ggplot() +
# geom_density(data = biov1, aes(x = lag), fill = "blue", color = "blue", alpha = .2) +
# geom_density(data = biov1, aes(x = lag2), fill = "red", color = "red", alpha = .2) +
# geom_density(data = biov1, aes(x = lag3), fill = "green", color = "green", alpha = .2) +
# geom_density(data = biov1, aes(x = lag4), fill = "yellow", color = "yellow", alpha = .2)

ggplot(biov1 %>%
           dplyr::filter(shift_vel_sign == 'pospos'), 
       aes(y=abs(lag3), fill = Param, color = Param))+
    geom_boxplot(alpha = .5)+
    scale_y_log10()+
    facet_wrap(.~Type, scales = "free")

```



# Load genetic data

```{r gen_data, eval=FALSE}

# Genetic diversity
mt <- fread(here::here('Data/mtdna.csv'))
mt$spp <- Clean_Names(mt$spp)

ms <- fread(here::here('Data/msat.csv'))
ms$spp <- Clean_Names(ms$spp)

gen_d <- fread(here::here('Data/Deposited_data_genetic_diversity_dekort2021.csv'))
gen_d$spp <-  gsub("_"," ",gen_d$Species)
gen_d$spp <- Clean_Names(gen_d$spp)

gen_lf <- read_excel(here::here("Data/MacroPopGen_Database_final-v0.2.xlsx"), sheet = "MacroPopGen_Database")
gen_lf$spp <-  gsub("_"," ",gen_lf$G_s)
gen_lf$spp <- Clean_Names(gen_lf$spp)

# Phenotypic rates
pheno <- fread(here::here('Data/PROCEEDv6_RatesDB.csv'))
pheno$spp <- Clean_Names(pheno$sp_ncbi)

length(unique(gen_d$spp))
length(unique(gen_lf$spp))
length(unique(mt$spp))
length(unique(ms$spp))
length(unique(pheno$spp))

```

## Fix names at the genetic data

```{r fixnames, eval=FALSE}

# Species to find
mycols <- c("species","scientificName","kingdom","phylum","class","order","family","db_code")
splist <- data.frame(matrix(ncol = length(mycols), nrow = length(unique(c(gen_sps,phen_sps)))))
names(splist) <- mycols

splist$reported_name_fixed <- unique(c(gen_sps,phen_sps))
tofind_ <- splist[which(is.na(splist$scientificName)),]
tofind_ <- unique(tofind_$reported_name_fixed)

tofind <- data.frame(matrix(nrow = length(tofind_), ncol = 8))
names(tofind) = c("scientificName", "kingdom", "phylum", "class", "order", "family", "db", "db_code")

tofind <- data.frame(species = tofind_, tofind)

tofind <- tofind %>%
    mutate(across(everything(), as.character))

# retrieve sp names
sp_names_found <- Find_Sci_Names(spnames = tofind$species)

# Feed
tofind <- tofind %>% 
    rows_patch(sp_names_found, 
               by = "species")

gc()

## Add found species names to the splist
all(tofind$species %in% splist$reported_name_fixed)

if(any(is.na(tofind$scientificName))){
    found <- tofind[-which(is.na(tofind$scientificName)),]
} else {
    found = tofind
}

for(i in 1:length(found$species)){
    tofill <- unique(which(splist$reported_name_fixed == found$species[i]))
    splist$scientificName[tofill] <- found$scientificName[i]
    splist$kingdom[tofill] <- found$kingdom[i]
    splist$phylum[tofill] <-found$phylum[i]
    splist$class[tofill] <- found$class[i]
    splist$order[tofill] <- found$order[i]
    splist$family[tofill] <- found$family[i]
    splist$db[tofill] <- found$db[i]
    splist$db_code[tofill] <- found$db_code[i]
}

splist$spp = splist$reported_name_fixed

# Keep original names for those species we could not find a name at GBIF
pos <- which(is.na(splist$scientificName))
splist$scientificName[pos] <- splist$spp[pos]

any(!mt$spp %in% gen_sps)
any(!mt$spp %in% splist$spp)

any(!ms$spp %in% gen_sps)
any(!ms$spp %in% splist$spp)

any(!gen_d$spp %in% gen_sps)
any(!gen_d$spp %in% splist$spp)

any(!gen_lf$spp %in% gen_sps)
any(!gen_lf$spp %in% splist$spp)

for(i in 1:nrow(mt)){
    mt$spp_new[i] <- splist$scientificName[which(splist$spp == mt$spp[i])]
}
for(i in 1:nrow(ms)){
    ms$spp_new[i] <- splist$scientificName[which(splist$spp == ms$spp[i])]
}
for(i in 1:nrow(gen_d)){
    gen_d$spp_new[i] <- splist$scientificName[which(splist$spp == gen_d$spp[i])]
}
for(i in 1:nrow(gen_lf)){
    gen_lf$spp_new[i] <- splist$scientificName[which(splist$spp == gen_lf$spp[i])]
}
for(i in 1:nrow(gen_lf)){
    pheno$spp_new[i] <- splist$scientificName[which(splist$spp == pheno$spp[i])]
}

write.csv(mt, 
          here::here('Data/mtdna_harmo.csv'), row.names = FALSE)
write.csv(ms, 
          here::here('Data/msat_harmo.csv'), row.names = FALSE)
write.csv(gen_d, 
          here::here('Data/Deposited_data_genetic_diversity_dekort2021_harmo.csv'), row.names = FALSE)
write.csv(gen_lf, 
          here::here('Data/MacroPopGen_Database_final_areas_Lawrence_Fraser_2020_harmo.csv'), row.names = FALSE)
write.csv(pheno, 
          here::here('Data/PROCEEDv6_RatesDB_harmo.csv'), row.names = FALSE)

```

# Load genetic data with harmonized taxonomies 

```{r load_harmonized_gen}

mt <- fread(here::here('Data/mtdna_harmo.csv'))
ms <- fread(here::here('Data/msat_harmo.csv'))
gen_d <- fread(here::here('Data/Deposited_data_genetic_diversity_dekort2021_harmo.csv'))
gen_d$spp <-  gsub("_"," ",gen_d$Species)
gen_lf <- fread(here::here('Data/MacroPopGen_Database_final_areas_Lawrence_Fraser_2020_harmo.csv'))
gen_lf$spp <-  gsub("_"," ",gen_lf$G_s)
pheno <- fread(here::here('Data/PROCEEDv6_RatesDB_harmo.csv'))
pheno$spp <-  gsub("_"," ",pheno$sp_ncbi)

```

## Stats

```{r}

# Gen
gen_sps <- unique(c(mt$spp, ms$spp, gen_d$spp, gen_lf$spp))

bsi_v1 <- intersect(biov1$spp, gen_sps)

# Break down
tot1 <- data.frame(N_species_with_gen_data = length(gen_sps),
                   v1_matches = length(bsi_v1))
tot1

# phenotypic
phen_sps <- unique(pheno$spp)

bsj_v1 <- intersect(biov1$spp, phen_sps)

# Break down
tot2 <- data.frame(N_species_with_pheno_data = length(phen_sps),
                   v1_matches = length(bsj_v1))
tot2

```

## Explore distribution

```{r}

# Malin
ggplot(mt, aes(He)) +
    geom_histogram() +
    ggtitle("Malin's Mitochondrial")

ggplot(ms, aes(He)) +
    geom_histogram() +
    ggtitle("Malin's Microsatellite")

# De Kort
ggplot(gen_d, aes(GDp)) +
    geom_histogram() +
    ggtitle("De Kort AFLP & Microsatellite")
# Why De Kort data has the two peaks? It is due to the marker type
ggplot(gen_d, aes(GDp)) +
    geom_histogram()+
    facet_grid(.~Marker)
# But if we look at the normalized GDp, the problem is solved. Therefore, we should be using the normalized version of GDp from DeKort
ggplot(gen_d, aes(GDp_norm)) +
    geom_histogram()

# Lawrence & Fraser
ggplot(gen_lf, aes(as.numeric(He))) +
    geom_histogram()+
    ggtitle("Lawrence & Fraser Microsatellite")

```

# Create a single genetic diversity dataset

```{r}

# vec<-rnorm(100)+abs(min(vec))
# vec<-vec+min(vec)
# 
# hist(vec)
# hist(transform01(vec))

# Malin
mt_sub <- mt  %>%
    mutate(long = as.numeric(lon),
           lat = as.numeric(lat),
           N = as.numeric(n), # Sample size
           Marker = "Mitochondrial")  %>%
    dplyr::filter(!is.na(He) | !He==0) %>%
    dplyr::select(spp, He, long, lat, Marker)
mt_sub$Source = "Malin"
mt_sub$Het_type = "He"

ms_sub <- ms %>%
    mutate(long = as.numeric(lon),
           lat = as.numeric(lat),
           N = as.numeric(n), # Sample size
           Marker = "Microsatellite")  %>%
    dplyr::filter(!is.na(He) | !He==0) %>%
    dplyr::select(spp, He, long, lat, Marker)
ms_sub$Source = "Malin"
ms_sub$Het_type = "He"

###################
# De Kort

# Jonathan R.:
# co-dominant refers to the fact that you can distinguish homozigous and heterozygous individuals, such as microsatellites
# dominant means that you either have the detection of the marker or not. but you don’t know if the individual is heterozygous or not, such as AFLP
# restriction enzymes are proteins that you use to cut the genomes into fragments. there are used for AFLP and allozymes, so not sure

# De Kort code for Marker:
# CD = Co-dominant
# D = Dominant
# ENZ = Enzime

gen_d_sub <- gen_d %>%
    mutate(long = as.numeric(LONGITUDE),
           lat = as.numeric(LATITUDE),
           spp = spp_new, 
           He = as.numeric(GDp),
           N = as.numeric(SampleSize), # Sample size
           Marker = case_when(
               Marker=="CD" ~ "Microsatellite",
               Marker=="D" ~ "AFLP",
               Marker=="ENZ" ~ "AFLP",
               TRUE ~ as.character(Marker))) %>%
    dplyr::filter(!is.na(He) | !He==0) %>%
    dplyr::filter(!(Marker == "AFLP" & He > .5)) %>% # AFLPs > 0.5 are errors
    dplyr::select(spp, He, long, lat, Marker)
gen_d_sub$Source = "De Kort"
gen_d_sub$Het_type = "He"

# How DeKort calculated the normalized version of GDp
# gen_d_sub$He_harm <- NA
# 
# m <- unique(gen_d_sub$Marker)
# for(i in 1:length(m)){
# pos <- which(gen_d_sub$Marker == m[i])
# gen_d_sub$He_harm[pos] <-  scale(gen_d_sub$He[pos])
# }
# gen_d_sub$He_harm <-  (gen_d_sub$He_harm - min(gen_d_sub$He_harm))/(max(gen_d_sub$He_harm)-min(gen_d_sub$He_harm))

###################
# Lawrence & Fraser
pos <- which(is.na(gen_lf$He))
gen_lf$He_new <- gen_lf$He
gen_lf$He_new[pos] <- gen_lf$Ho[pos]
gen_lf$Het_type = "He"
gen_lf$Het_type[pos] = "Ho"

gen_lf_sub <- gen_lf %>%
    mutate(long = as.numeric(Long),
           lat = as.numeric(Lat),
           He = as.numeric(He_new),
           spp = spp_new,
           N = as.numeric(n), # Sample size
           Marker = "Microsatellite") %>% 
    dplyr::filter(!is.na(He) | !He==0) %>%
    dplyr::select(spp, He, long, lat, Marker, Het_type) 
gen_lf_sub$Source = "Lawrence & Fraser"

###################
# Group all
gen_div <- rbind(mt_sub %>% dplyr::select(names(mt_sub)), 
                 ms_sub %>% dplyr::select(names(mt_sub)),
                 gen_d_sub %>% dplyr::select(names(mt_sub)),
                 gen_lf_sub %>% dplyr::select(names(mt_sub)))
gen_div <- na.omit(gen_div)

gen_div$spp <- as.factor(gen_div$spp)
gen_div$Marker <- as.factor(gen_div$Marker)
gen_div$Source <- as.factor(gen_div$Source)

```

# Harmonize genetic data

- He_harm = transformation used in DeKort et al 2021 (Nature Communications). Makes markers range from 0-1.
1) scale values at each marker type
2) standardize (x - min)/(max-min)
- He_harm2 = transformation suggested by Malin. Applies the function transform01 before a logit transformation. 
1) transform01 >> The only intent of this function is to move slightly inward the values of He so that it allows applying logit transformation.
2) apply logit transformation
- He_harm3 = He_harm2 centered in zero. 

```{r}

gen_div$He_harm <- NA
gen_div$He_harm2 <- NA
gen_div$He_harm3 <- NA

m <- unique(gen_div$Marker)
for(i in 1:length(m)){
    # Select marker type
    pos <- which(gen_div$Marker == m[i])
    # Apply DeKort transformation
    gen_div$He_harm[pos] <- deKort_trans(gen_div$He[pos])
    # Apply logit transformation
    gen_div$He_harm2[pos] <- logit_trans(gen_div$He[pos])
    # Centered in zero
    gen_div$He_harm3[pos] <- scale(gen_div$He_harm2[pos])
}

# Filter the species in Bioshifts
gen_div <- gen_div %>% dplyr::filter(spp %in% unique(biov1$spp))

# remove outliers
gen_metrics <- c("He_harm","He_harm2","He_harm3")

for (i in 1:length(gen_metrics)){
    my_gem <- as.numeric(data.frame(gen_div %>% select(gen_metrics[i]))[,1])
    
    quartiles <- quantile(my_gem, probs=c(.05, .95), na.rm = FALSE)
    IQR <- IQR(my_gem)
    
    Lower <- quartiles[1] - 1.5*IQR
    Upper <- quartiles[2] + 1.5*IQR 
    
    gen_div <- subset(gen_div, my_gem > Lower & my_gem < Upper)
}


```

## Avg by location 
We averaged genetic diversity by location (same species, in the same XY coordinates)
```{r}
# How many species with >1 gen div measurements at the same site?
# Check how many obs per site and marker type exist
N_obs <- gen_div %>%
    group_by(spp, long, lat) %>%
    tally() %>%
    dplyr::filter(n>1)

DT::datatable(N_obs)

# Is there any species with >1 marker type in the same location?
gen_div %>%
    group_by(spp, long, lat, Marker) %>%
    dplyr::summarise(N = length(unique(spp))) %>%
    dplyr::filter(N>1)
# No, there are not

# Is there any species with >1 Source type in the same location?
gen_div %>%
    group_by(spp, long, lat, Source) %>%
    dplyr::summarise(N = length(unique(spp))) %>%
    dplyr::filter(N>1)
# No, there are not

# Avg by site and species
gen_div <- gen_div %>%
    group_by(spp, long, lat) %>%
    dplyr::summarise(He = median(He),
                     He_harm = median(He_harm),
                     He_harm2 = median(He_harm2),
                     He_harm3 = median(He_harm3),
                     Source = paste(Source, sep = ", "),
                     Marker = paste(Marker, sep = ", "))

unique(gen_div$Source)
unique(gen_div$Marker)

```

## Which species?

```{r fig.height=7, fig.width=5}

gen_div <- merge(gen_div, 
                 splist[,-7],
                 by.x="spp",
                 by.y="scientificName",
                 all.x = T)

toplot_ <- gen_div %>%
    group_by(Group) %>%
    dplyr::summarise(Obs = length(spp),
                     Species = length(unique(spp))) %>%
    gather(var, N, -c(Group))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N, 
                  hjust = ifelse(N<max(N),-.1,1.1)), vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()+
    facet_wrap(.~var, scales = "free", ncol=1)

```

## Distribution of He

```{r message=FALSE}
# correlations
ggpairs(gen_div,
        columns = c("He","He_harm","He_harm2","He_harm3"))


ggpairs(gen_div,
        columns = c("He","He_harm","He_harm2","He_harm3"),      
        aes(color = Marker,  
            alpha = 0.5))     

ggpairs(gen_div,
        columns = c("He","He_harm","He_harm2","He_harm3"),      
        aes(color = Source,  
            alpha = 0.5))
```


```{r message=FALSE, fig.width=10,fig.height=3}

to_plot <- gen_div %>%
    select(Group,He_harm,He_harm2,He_harm3) %>%
    filter(Group %in% c("Mammal","Fish","Plant", "Bird")) %>%
    gather(metric, value, -Group)

ggplot(to_plot, aes(value, fill = Group, color = Group))+
    geom_density(alpha = .3)+
    scale_color_viridis_d()+
    scale_fill_viridis_d()+
    facet_wrap(.~metric, scales = "free")


```

## Spatial pattern
```{r fig.width=10, fig.height=5}

mundi <- terra::vect(rnaturalearth::ne_coastline(scale = 110, returnclass = "sp"))

# get vect gen div
vect_div <- terra::vect(gen_div,geom=c("long","lat"),crs=crs(mundi))

# get vect centroids bioshifts
vect_biov1 <- terra::vect(biov1 %>%
                              filter(spp %in% gen_div$spp),
                          geom=c("centroid.x","centroid.y"),crs=crs(mundi))

# get raster bioshifts shp files study areas
get_raster_bioshifts = "NO"
if(get_raster_bioshifts=="YES"){
    my_ext = terra::ext(mundi)
    my_crs = crs(mundi)
    rast_biov1 <- terra::rast(my_ext, crs = my_crs, res = 0.5)
    values(rast_biov1) <- 0
    
    fgdb <- "C:/Users/brunn/NextCloud/Bioshifts/Study_Areas.gdb"
    fc_list <- ogrListLayers(fgdb)
    fc_list <- fc_list[which(fc_list %in% unique(vect_biov1$ID))]
    
    for(i in 1:length(fc_list)){ cat("\r",i,"from",length(fc_list))
        tmp = terra::vect(sf::st_read(fgdb, layer=fc_list[i]))
        tmp = terra::cells(rast_biov1,tmp)
        tmp_cell = tmp[,2]
        tmp_vals = rast_biov1[tmp_cell][,1]
        rast_biov1[tmp_cell] = tmp_vals+1
    }
    names(rast_biov1) <- "SA"
    rast_biov1[rast_biov1==0] <- NA
    
    writeRaster(rast_biov1, "Data/raster_bioshifts_SA.tif")
    
} else {
    rast_biov1 <- terra::rast("Data/raster_bioshifts_SA.tif")
}

values_biov1 <- na.omit(values(rast_biov1))

ggplot()+
    ggtitle("Genetic diversity data")+
    geom_spatvector(data=mundi)+
    geom_spatvector(data=vect_div,aes(color = Marker))+
    theme_blank()

ggplot()+
    ggtitle("Genetic diversity data")+
    geom_spatvector(data=mundi)+
    geom_spatvector(data=vect_div,aes(color = Source))+
    theme_blank()

ggplot()+
    ggtitle("Bioshifts data \n(Centroid of study areas)")+
    geom_spatvector(data=mundi)+
    geom_spatvector(data=vect_biov1, aes(color = ECO))+
    theme_blank()

ggplot()+
    ggtitle("Bioshifts data \n(Overlap study areas)")+
    geom_spatraster(data=rast_biov1)+
    scale_fill_whitebox_b(
        palette = "muted",
        na.value = "white",
        breaks = seq(min(values_biov1), max(values_biov1), 1))+
    geom_spatvector(data=vect_biov1, aes(color = ECO))+
    geom_spatvector(data=mundi)+
    theme_blank()


```



# Phenotypic rates
```{r message=FALSE}
pheno_rate <- pheno %>%
    mutate(long = as.numeric(sample1.longitude),
           lat = as.numeric(sample1.latitude),
           trait1 = as.numeric(mean1),
           trait2 = as.numeric(mean2),
           spp = spp_new,
           Trait = trait_type, 
           Years = as.numeric(years),
           Gen = as.numeric(generations)) %>% 
    dplyr::select(spp, long, lat, Trait, trait1, trait2, Years, Gen) %>%
    dplyr::filter(!is.na(trait1) & !is.na(trait2))
pheno_rate$Source = "PROCEED"
```

## Calculate rate & harmozine

```{r message=FALSE}
for(i in 1:nrow(pheno_rate)){
    sppi <- pheno_rate$spp[i]
    traiti <- pheno_rate$Trait[i]
    pos_spp_trait_i <- which(pheno_rate$spp == sppi & pheno_rate$Trait == traiti)
    pheno_rate$PR1[i] <- (pheno_rate$trait2[i] - pheno_rate$trait1[i]) / mean(c(pheno_rate$trait2[pos_spp_trait_i], pheno_rate$trait1[pos_spp_trait_i])) * pheno_rate$Years[i]
    pheno_rate$PR2[i] <- (pheno_rate$trait2[i] - pheno_rate$trait1[i]) / sqrt(mean(c(pheno_rate$trait2[pos_spp_trait_i], pheno_rate$trait1[pos_spp_trait_i])))^2 * pheno_rate$Years[i]
}

pheno_rate$PR_harm <- NA

m <- unique(pheno_rate$Trait)
for(i in 1:length(m)){
    pos <- which(pheno_rate$Trait == m[i])
    # DeKort
    p <- scale(pheno_rate$PR1[pos])
    p <- (p - min(p, na.rm = T))/(max(p, na.rm = T)-min(p, na.rm = T))
    pheno_rate$PR_harm[pos] <-  p
}

# gen_div %>%
# group_by(spp) %>%
# tally() %>%
# summary()

```


## Distribution of pheno rates
```{r message=FALSE}
# Look into phenotic rates
ggplot(pheno_rate, aes(PR1)) +
    geom_histogram() +
    facet_wrap(Trait~., scales = "free") +
    ggtitle("Raw data")

ggplot(pheno_rate, aes(PR_harm)) +
    geom_histogram() +
    facet_wrap(Trait~., scales = "free") +
    ggtitle("Harmonization following De Kort")
```



# Merge genetic and bioshifts data
We tried several methods for merging genetic data and bioshifts. We decided that Method 5 is the best one.

# Method 1 - Normal merge
This is not the correct way to merge because it creates multiple false duplicates.
```{r eval=FALSE, include=FALSE}

gen_data_v1 <- append(list(gen_div), 
                      list(biov1)) %>% reduce(full_join, by='spp')
gen_data_v1 <- gen_data_v1 %>% 
    dplyr::filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1$spp))

# Use only plus plus or minus minus
table(gen_data_v1$shift_vel_sign)

gen_data_v1 <- gen_data_v1 %>% dplyr::filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

gen_data_v1
# N species
length(unique(gen_data_v1$spp))

# N obs per species
# View(table(gen_data_v1$spp))

```

## Which species?
```{r eval=FALSE, include=FALSE}
toplot_ <- gen_data_v1 %>%
    group_by(Class) %>%
    dplyr::summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Class, y = N))+
    ggtitle("N species by Class")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=4}
gen_data_v1 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=5}

gen_data_v1 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

The merge we made considers all possible combinations of He and Shifts for each species. Therefore, there are lots of pseudo-replicates. Explore average difference in genetic diversity measures per species.

```{r eval=FALSE, include=FALSE}

DT::datatable(gen_div %>% 
                  group_by(spp) %>%
                  dplyr::summarise(mean = mean(He_harm2),
                                   sd = sd(He_harm2),
                                   N = length(spp)))

# For some species we can get quite large variance in He. For example Fistularia commersonii:
gen_div %>% 
    dplyr::filter(spp=="Fistularia commersonii")


```

# Method 2 - Median He per species
This ignores spatial proximity of He values to range shift values.
```{r eval=FALSE, include=FALSE}

gen_div_avg <- gen_div %>%
    group_by(spp) %>%
    dplyr::summarise(N = length(spp),
                     He = median(He_harm)) 

#merge all data frames in list
gen_data_v1_avg <- append(list(gen_div_avg), list(biov1)) %>% reduce(full_join, by='spp')
gen_data_v1_avg <- gen_data_v1_avg %>% dplyr::filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_avg$spp))

# Use only plus plus or minus minus
table(gen_data_v1_avg$shift_vel_sign)

gen_data_v1_avg <- gen_data_v1_avg %>% dplyr::filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_avg$spp))

```

## Which species?

```{r eval=FALSE, include=FALSE}
toplot_ <- gen_data_v1_avg %>%
    group_by(Group) %>%
    dplyr::summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=4}
gen_data_v1_avg %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_avg %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=5}

gen_data_v1_avg %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_avg %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

# Method 3 - Median He per Param & species

Use lower latitude TE and higher latitude as LE (North Hemisphere).  
Although this collects He values based on latitude, the proximity of He to the study areas where range shifts were observed is ignored.

```{r eval=FALSE, include=FALSE, error=FALSE}

gen_div_avg_2 <- gen_div %>%
    dplyr::group_by(spp) %>%
    dplyr::summarise(
        LE = ifelse(lat>0, # weighted by latitude
                    weighted.mean(He_harm, lat),
                    weighted.mean(He_harm, 1/lat)), 
        O = median(He_harm),
        TE = ifelse(lat>0, # weighted by latitude
                    weighted.mean(He_harm, 1/lat),
                    weighted.mean(He_harm, lat))) %>%
    gather(Param, He, -spp)

#merge all data frames in list
gen_data_v1_avg_2 <- append(list(gen_div_avg_2), list(biov1)) %>% reduce(full_join, by=c("spp","Param"))
gen_data_v1_avg_2 <- gen_data_v1_avg_2 %>% dplyr::filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_avg_2$spp))

# Use only plus plus or minus minus
table(gen_data_v1_avg_2$shift_vel_sign)

gen_data_v1_avg_2 <- gen_data_v1_avg_2 %>% dplyr::filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_avg_2$spp))

```

## Which species?

```{r eval=FALSE, include=FALSE}
toplot_ <- gen_data_v1_avg_2 %>%
    group_by(Group) %>%
    dplyr::summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=4}

gen_data_v1_avg_2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_avg_2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

```

## Per group

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=5}

gen_data_v1_avg_2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_avg_2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

# Method 4 - Based on distance

Calculate distance weighted mean He for each species in the bioshifts database.  
This method collects He values more close to the centroid of the study area.

```{r eval=FALSE, include=FALSE, message=FALSE}

# sp list
m <- unique(biov1$spp[which(biov1$spp %in% gen_div$spp)])

gen_data_v1_dist <- list()

# sp2go="Alces alces"
# sp2go="Sylvia atricapilla"
for(i in 1:length(m)){ cat(i, "from", length(m), "\r")
    sp2go <- m[i]
    sub_gen <- subset(gen_div, spp == sp2go)
    sub_gen_v <- terra::vect(sub_gen, 
                             geom=c("long", "lat"), 
                             crs = "+proj=longlat +datum=WGS84 +no_defs")
    
    sub_bio <- subset(biov1, spp == sp2go)
    areas <- unique(sub_bio$ID)
    
    gen_data_v1_dist_tmp <- list()
    
    for(j in 1:length(areas)){
        area2go <- areas[j]
        sub_bio_area_j <- subset(sub_bio, ID == area2go)
        sub_bio_area_j_v <- terra::vect(unique(sub_bio_area_j[,c("centroid.x", "centroid.y")]), 
                                        geom=c("centroid.x", "centroid.y"), 
                                        crs = "+proj=longlat +datum=WGS84 +no_defs")
        
        dists <- terra::distance(sub_bio_area_j_v, sub_gen_v)[1,]
        
        # the closest
        dists_order <- order(dists)[1]
        # sub_gen <- sub_gen[dists_order,]
        
        # gen_data_v1_dist_tmp[[j]] <- append(list(sub_gen), list(sub_bio_area_j)) %>% 
        # reduce(full_join, by="spp")
        
        # weighted distance
        wtd_he <- weighted.mean(sub_gen$He_harm, 1/dists) # inverse of the distance
        sub_gen_tmp <- sub_gen[dists_order,]
        sub_gen_tmp$He <- wtd_he
        
        gen_data_v1_dist_tmp[[j]] <- append(list(sub_gen_tmp), list(sub_bio_area_j)) %>% 
            reduce(full_join, by="spp")
    }
    
    gen_data_v1_dist[[i]] <- rbindlist(gen_data_v1_dist_tmp)
    
}

gen_data_v1_dist <- rbindlist(gen_data_v1_dist)

gen_data_v1_dist <- gen_data_v1_dist %>% dplyr::filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_dist$spp))

# Use only plus plus or minus minus
table(gen_data_v1_dist$shift_vel_sign)

gen_data_v1_dist <- gen_data_v1_dist %>% dplyr::filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_dist$spp))

```

## Which species?

```{r eval=FALSE, include=FALSE}
toplot_ <- gen_data_v1_dist %>%
    group_by(Group) %>%
    dplyr::summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=4}
gen_data_v1_dist %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_dist %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r eval=FALSE, include=FALSE, fig.width=10, fig.height=5}

gen_data_v1_dist %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_dist %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

```


# Method 5 - Based on location

Get the closest He to the centroid (for O), to the max latitude (for LE, at the North hemisphere), and to the min latitude (for TE, at the North hemisphere) of the study area.

```{r message=FALSE}

# sp list
m <- unique(biov1$spp[which(biov1$spp %in% gen_div$spp)])

gen_data_v1_dist2 <- data.frame()

# sp2go="Alces alces"
# sp2go="Sylvia atricapilla"
for(i in 1:length(m)){ cat(i, "from", length(m), "\r")
    sp2go <- m[i]
    
    # subset genetic data
    sub_gen <- subset(gen_div[,1:9], spp == sp2go)
    sub_gen_v <- terra::vect(sub_gen, 
                             geom=c("long", "lat"), 
                             crs = "+proj=longlat +datum=WGS84 +no_defs")
    
    # subset bioshifts data
    sub_bio <- subset(biov1, spp == sp2go)
    
    # study areas for species i
    areas <- unique(sub_bio$ID)
    
    gen_data_v1_dist2_tmp <- data.frame()
    
    # get genetic data for species i in each study area
    for(j in 1:nrow(sub_bio)){
        
        # bioshifts in study area j and range pos r
        sub_bio_area_j_tmp <- sub_bio[j,]
        
        range_pos2go <- sub_bio_area_j_tmp$Param
        
        coord_names <- NA
        coord_names <- if(range_pos2go == "O"){ c("centroid.x", "centroid.y") } else {coord_names}
        coord_names <- if(range_pos2go == "LE" & sub_bio_area_j_tmp$centroid.y > 0){ c("max_lat.x", "max_lat.y") } else {coord_names}
        coord_names <- if(range_pos2go == "LE" & sub_bio_area_j_tmp$centroid.y < 0){ c("min_lat.x", "min_lat.y") } else {coord_names}
        coord_names <- if(range_pos2go == "TE" & sub_bio_area_j_tmp$centroid.y < 0){ c("max_lat.x", "max_lat.y") } else {coord_names}
        coord_names <- if(range_pos2go == "TE" & sub_bio_area_j_tmp$centroid.y > 0){ c("min_lat.x", "min_lat.y") } else {coord_names}
        
        sub_bio_area_j_v = sub_bio_area_j_tmp[,coord_names]
        names(sub_bio_area_j_v) <- c("long", "lat")
        
        sub_bio_area_j_v <- terra::vect(unique(sub_bio_area_j_v[,c("long", "lat")]), 
                                        geom=c("long", "lat"), 
                                        crs = "+proj=longlat +datum=WGS84 +no_defs")
        
        # plot(terra::vect(rbind(sub_gen[,c("long", "lat")], sub_bio_area_j_tmp[,c("long", "lat")]),
        #  geom=c("long", "lat"),
        #  crs = "+proj=longlat +datum=WGS84 +no_defs"))
        # plot(sub_bio_area_j_v, col = "red", add=T)
        
        dists <- terra::distance(sub_bio_area_j_v, sub_gen_v)[1,]
        dists_order <- order(dists)[1]
        min_dist <- dists[dists_order] # distance to closest
        
        # the closest
        sub_gen_tmp <- sub_gen[dists_order,]
        sub_gen_tmp$min_dist <- min_dist
        
        gen_data_v1_dist2_tmp <- rbind(gen_data_v1_dist2_tmp,
                                       append(list(sub_gen_tmp), list(sub_bio_area_j_tmp)) %>%
                                           reduce(full_join, by="spp"))
        
        # weighted distance
        # wtd_he <- weighted.mean(sub_gen$He_harm, 1/dists) # inverse of the distance
        # sub_gen_tmp <- sub_gen[dists_order,]
        # sub_gen_tmp$He <- wtd_he
        # 
        # gen_data_v1_dist2_tmp <- rbind(gen_data_v1_dist2_tmp,
        #                            append(list(sub_gen_tmp), list(sub_bio_area_j_tmp)) %>%
        # reduce(full_join, by="spp"))
        
    }
    
    gen_data_v1_dist2 <- rbind(gen_data_v1_dist2,
                               gen_data_v1_dist2_tmp)
    
}

gen_data_v1_dist2 <- gen_data_v1_dist2 %>% dplyr::filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_dist2$spp))

# Use only plus plus or minus minus
table(gen_data_v1_dist2$shift_vel_sign)

gen_data_v1_dist2 <- gen_data_v1_dist2 %>% dplyr::filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_dist2$spp))

```

## Which species?

```{r fig.width=5,fig.height=10}

toplot_ <- gen_data_v1_dist2 %>%
    group_by(Group) %>%
    dplyr::summarise(Species = length(unique(spp)),
                     Obs = length(spp))

toplot_ <- reshape2::melt(toplot_)

ggplot(toplot_, aes(x = Group, y = value))+
    geom_bar(stat="identity")+
    geom_text(aes(y=value, label=value), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()+
    facet_wrap(.~variable, scales = "free", ncol = 1)

```

## General pattern

```{r fig.width=10, fig.height=4}

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He_harm, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He harm)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He_harm, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He harm)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He_harm3, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He harm logit)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He_harm3, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He harm logit)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r fig.width=10, fig.height=5}

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He_harm, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He harm)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He_harm, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He harm)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "LAT") %>%
    ggplot(aes(x = He_harm3, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He harm logit)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_dist2 %>%
    dplyr::filter(Type == "ELE") %>%
    ggplot(aes(x = He_harm3, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He harm logit)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

```


# Playing with models

## Choose dataset

```{r message=FALSE}

# mydataset <- gen_data_v1_dist
mydataset <- gen_data_v1_dist2

```

## Latitude

```{r message=FALSE}
mydatatogo <- get_data_to_go(mydataset, Type = "LAT")
```

## Remove outliers in shift

```{r message=FALSE}
summary(mydataset$SHIFT)
hist(mydataset$SHIFT)
thr <- quantile(mydataset$SHIFT,c(.05,0.95))
thr

rem <- which((mydataset$SHIFT > thr[1]) & (mydataset$SHIFT < thr[2]))
if(any(rem)){
    mydataset <- mydataset[rem,]
}
summary(mydataset$SHIFT)
hist(mydataset$SHIFT)
# hist(log(range01(mydataset$SHIFT)))
mydatatogo$SHIFT <- logit(transform01(range01(mydatatogo$SHIFT)))

##############
# Lags look ok

# remove outliers in lag
# summary(mydataset$lag)
hist(mydataset$lag)
# thr <- quantile(mydataset$lag,c(.01,0.99))
# thr
# 
# rem <- which((mydataset$lag > thr[1]) & (mydataset$lag < thr[2]))
# if(any(rem)){
#     mydataset <- mydataset[rem,]
# }
# summary(mydataset$lag)
# hist(mydataset$lag)
```

### Lag
#### He harm
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = lag, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(lag ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He*Group*Param + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")

# Plot fixed effects of m1
# plot_model(m2, type = "pred", terms = c("He","Param","Group"))
m_ef <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef)[c(3,6,7)] <- c("Coeff","LCI","UCI")

ggplot(m_ef, aes(Coeff, Param, color = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = LCI, xmax = UCI))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")

# Plot random effects
eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
tmp <- lapply(eff_plot$level, function(x) {
    strsplit(x,":")[[1]]
})
eff_plot$Group <- sapply(tmp, function(x) x[1])
eff_plot$Param <- sapply(tmp, function(x) x[2])

ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
    facet_wrap(.~Group)+
    xlab("Random effect")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Different to fixed effects of m2

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

#### He harm 3
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm3

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = lag, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(lag ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He*Group*Param + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")

# Plot fixed effects of m1
# plot_model(m2, type = "pred", terms = c("He","Param","Group"))
m_ef <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef)[c(3,6,7)] <- c("Coeff","LCI","UCI")

ggplot(m_ef, aes(Coeff, Param, color = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = LCI, xmax = UCI))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")

# Plot random effects
eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
tmp <- lapply(eff_plot$level, function(x) {
    strsplit(x,":")[[1]]
})
eff_plot$Group <- sapply(tmp, function(x) x[1])
eff_plot$Param <- sapply(tmp, function(x) x[2])

ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
    facet_wrap(.~Group)+
    xlab("Random effect")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Different to fixed effects of m2

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

### Shift 
#### He harm
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")

p.vel <- ggplot(data = mydatatogo,
                aes(x = velocity, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Temperature velocity")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(SHIFT ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*Group*Param*velocity) + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*velocity) +
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param)+
               (velocity|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")
confint(m1, parm = "velocity")

# Plot fixed effects of m1
# plot_model(m1, type = "pred", terms = c("He","Param","Group"))
m_ef1 <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef1)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef2 <- data.frame(emtrends(m1, ~ Param*Group, var="velocity"))
names(m_ef2)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef <- as.data.frame(rbind(data.frame(x= "He", m_ef1),
                            data.frame(x= "velocity", m_ef2)))

ggplot(m_ef, aes(Coeff, Param, color = x))+
    geom_pointrange(aes(xmin = LCI, xmax = UCI),position = position_dodge2(w = 0.75))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

# Partial residuals from fixed effects
tt=as.data.frame(effect("Group:Param:velocity", partial.residuals=T, m1))

p.vel + 
    geom_line(data=tt, aes(x=velocity, y=fit)) +
    ylab("Partial residuals (Velocity)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")
confint(m2, parm = "velocity")

# Plot random effects
# Does not work because the functions doesnt know how to deal with the multiple random slopes
# eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
# eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
# tmp <- lapply(eff_plot$level, function(x) {
#     strsplit(x,":")[[1]]
# })
# eff_plot$Group <- sapply(tmp, function(x) x[1])
# eff_plot$Param <- sapply(tmp, function(x) x[2])
# 
# ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
#     geom_point()+
#     geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
#     facet_wrap(term~Group)+
#     xlab("Random effect")+
#     geom_vline(xintercept = 0, linetype = "dashed")

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

#### He harm 3
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm3

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")

p.vel <- ggplot(data = mydatatogo,
                aes(x = velocity, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Temperature velocity")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(SHIFT ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*Group*Param*velocity) + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*velocity) +
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param)+
               (velocity|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")
confint(m1, parm = "velocity")

# Plot fixed effects of m1
# plot_model(m1, type = "pred", terms = c("He","Param","Group"))
m_ef1 <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef1)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef2 <- data.frame(emtrends(m1, ~ Param*Group, var="velocity"))
names(m_ef2)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef <- as.data.frame(rbind(data.frame(x= "He", m_ef1),
                            data.frame(x= "velocity", m_ef2)))

ggplot(m_ef, aes(Coeff, Param, color = x))+
    geom_pointrange(aes(xmin = LCI, xmax = UCI),position = position_dodge2(w = 0.75))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

# Partial residuals from fixed effects
tt=as.data.frame(effect("Group:Param:velocity", partial.residuals=T, m1))

p.vel + 
    geom_line(data=tt, aes(x=velocity, y=fit)) +
    ylab("Partial residuals (Velocity)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")
confint(m2, parm = "velocity")

# Plot random effects
# Does not work because the functions doesnt know how to deal with the multiple random slopes
# eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
# eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
# tmp <- lapply(eff_plot$level, function(x) {
#     strsplit(x,":")[[1]]
# })
# eff_plot$Group <- sapply(tmp, function(x) x[1])
# eff_plot$Param <- sapply(tmp, function(x) x[2])
# 
# ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
#     geom_point()+
#     geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
#     facet_wrap(term~Group)+
#     xlab("Random effect")+
#     geom_vline(xintercept = 0, linetype = "dashed")

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)




```




## Elevation

```{r message=FALSE}

mydatatogo <- get_data_to_go(mydataset, Type = "ELE")

```

## Remove outliers in shift

```{r message=FALSE}
summary(mydataset$SHIFT)
hist(mydataset$SHIFT)
thr <- quantile(mydataset$SHIFT,c(.05,0.95))
thr

rem <- which((mydataset$SHIFT > thr[1]) & (mydataset$SHIFT < thr[2]))
if(any(rem)){
    mydataset <- mydataset[rem,]
}
summary(mydataset$SHIFT)
hist(mydataset$SHIFT)
# hist(log(range01(mydataset$SHIFT)))
mydatatogo$SHIFT <- logit(transform01(range01(mydatatogo$SHIFT)))

##############
# Lags look ok

# remove outliers in lag
# summary(mydataset$lag)
hist(mydataset$lag)
# thr <- quantile(mydataset$lag,c(.01,0.99))
# thr
# 
# rem <- which((mydataset$lag > thr[1]) & (mydataset$lag < thr[2]))
# if(any(rem)){
#     mydataset <- mydataset[rem,]
# }
# summary(mydataset$lag)
# hist(mydataset$lag)
```

### Lag
#### He harm
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = lag, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(lag ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He*Group*Param + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")

# Plot fixed effects of m1
# plot_model(m2, type = "pred", terms = c("He","Param","Group"))
m_ef <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef)[c(3,6,7)] <- c("Coeff","LCI","UCI")

ggplot(m_ef, aes(Coeff, Param, color = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = LCI, xmax = UCI))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")

# Plot random effects
eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
tmp <- lapply(eff_plot$level, function(x) {
    strsplit(x,":")[[1]]
})
eff_plot$Group <- sapply(tmp, function(x) x[1])
eff_plot$Param <- sapply(tmp, function(x) x[2])

ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
    facet_wrap(.~Group)+
    xlab("Random effect")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Different to fixed effects of m2

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

#### He harm 3
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm3

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = lag, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(lag ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He*Group*Param + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(lag ~ 
               # Fixed effects
               ## Ecology
               He + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")

# Plot fixed effects of m1
# plot_model(m2, type = "pred", terms = c("He","Param","Group"))
m_ef <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef)[c(3,6,7)] <- c("Coeff","LCI","UCI")

ggplot(m_ef, aes(Coeff, Param, color = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = LCI, xmax = UCI))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")

# Plot random effects
eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
tmp <- lapply(eff_plot$level, function(x) {
    strsplit(x,":")[[1]]
})
eff_plot$Group <- sapply(tmp, function(x) x[1])
eff_plot$Param <- sapply(tmp, function(x) x[2])

ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
    geom_point()+
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
    facet_wrap(.~Group)+
    xlab("Random effect")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Different to fixed effects of m2

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

### Shift 
#### He harm
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")

p.vel <- ggplot(data = mydatatogo,
                aes(x = velocity, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Temperature velocity")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(SHIFT ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*Group*Param*velocity) + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*velocity) +
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param)+
               (velocity|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")
confint(m1, parm = "velocity")

# Plot fixed effects of m1
# plot_model(m1, type = "pred", terms = c("He","Param","Group"))
m_ef1 <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef1)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef2 <- data.frame(emtrends(m1, ~ Param*Group, var="velocity"))
names(m_ef2)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef <- as.data.frame(rbind(data.frame(x= "He", m_ef1),
                            data.frame(x= "velocity", m_ef2)))

ggplot(m_ef, aes(Coeff, Param, color = x))+
    geom_pointrange(aes(xmin = LCI, xmax = UCI),position = position_dodge2(w = 0.75))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

# Partial residuals from fixed effects
tt=as.data.frame(effect("Group:Param:velocity", partial.residuals=T, m1))

p.vel + 
    geom_line(data=tt, aes(x=velocity, y=fit)) +
    ylab("Partial residuals (Velocity)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")
confint(m2, parm = "velocity")

# Plot random effects
# Does not work because the functions doesnt know how to deal with the multiple random slopes
# eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
# eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
# tmp <- lapply(eff_plot$level, function(x) {
#     strsplit(x,":")[[1]]
# })
# eff_plot$Group <- sapply(tmp, function(x) x[1])
# eff_plot$Param <- sapply(tmp, function(x) x[2])
# 
# ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
#     geom_point()+
#     geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
#     facet_wrap(term~Group)+
#     xlab("Random effect")+
#     geom_vline(xintercept = 0, linetype = "dashed")

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)



```

#### He harm 3
```{r }

# define the He metric to go
mydatatogo$He <- mydatatogo$He_harm3

p.gen <- ggplot(data = mydatatogo,
                aes(x = He, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")

p.vel <- ggplot(data = mydatatogo,
                aes(x = velocity, y = SHIFT, color = Group))+
    ggtitle("Latitude")+
    xlab("Temperature velocity")+
    ylab("Range shift (km/yr)") +
    geom_point(alpha = .5)+
    theme_bw()+ 
    facet_grid(Param ~ Group, scales = "free")


# Select which methodological variables to include in the model
m_lat_test <- lm(SHIFT ~ Method + START + DUR + ID.area + Sampling,
                 data = mydatatogo,
                 na.action = "na.fail")

test <- MuMIn::dredge(m_lat_test)
# View(subset(test, delta < 4))
# All variables are important

# Find the best model structure
m1 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*Group*Param*velocity) + 
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")
m2 <- lmer(SHIFT ~ 
               # Fixed effects
               ## Ecology
               (He*velocity) +
               ## Methods
               Method + START + DUR + ID.area + Sampling + Marker +
               # Random effects
               ## Taxonomy
               (1|spp) +
               ## Ecology
               (He|Group:Param)+
               (velocity|Group:Param),
           # weight by distance of genetic data to range shift data
           weights = dist_weight,
           # Other LME params
           control = lmerControl(optimizer ="Nelder_Mead"),
           REML=TRUE,
           data = mydatatogo,
           na.action = "na.fail")

model_list <- list(m1,m2)
m_metr <- lapply(model_list, function(x){
    tmp = data.frame(model = as.character(x@call[2]), BIC = BIC(x),AIC=AIC(x))
    r2s = MuMIn::r.squaredGLMM(x)
    cbind(tmp,r2s)
}
)
m_metr <- do.call(rbind,m_metr)
m_metr$DeltaAIC <- m_metr$AIC - min(m_metr$AIC)
m_metr[order(m_metr$BIC),]

# Check residuals of both models
plot(m1)
plot(m2)

###################
# Look at model 1:

summary(m1)
confint(m1, parm = "He")
confint(m1, parm = "velocity")

# Plot fixed effects of m1
# plot_model(m1, type = "pred", terms = c("He","Param","Group"))
m_ef1 <- data.frame(emtrends(m1, ~ Param*Group, var="He"))
names(m_ef1)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef2 <- data.frame(emtrends(m1, ~ Param*Group, var="velocity"))
names(m_ef2)[c(3,6,7)] <- c("Coeff","LCI","UCI")

m_ef <- as.data.frame(rbind(data.frame(x= "He", m_ef1),
                            data.frame(x= "velocity", m_ef2)))

ggplot(m_ef, aes(Coeff, Param, color = x))+
    geom_pointrange(aes(xmin = LCI, xmax = UCI),position = position_dodge2(w = 0.75))+
    facet_wrap(.~Group, scales = "free")+
    xlab("Std. Coefficient")+
    geom_vline(xintercept = 0, linetype = "dashed")

# Partial residuals from fixed effects
tt=as.data.frame(effect("He:Group:Param", partial.residuals=T, m1))

p.gen + 
    geom_line(data=tt, aes(x=He, y=fit)) +
    ylab("Partial residuals (He)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

# Partial residuals from fixed effects
tt=as.data.frame(effect("Group:Param:velocity", partial.residuals=T, m1))

p.vel + 
    geom_line(data=tt, aes(x=velocity, y=fit)) +
    ylab("Partial residuals (Velocity)")+
    geom_ribbon(data = tt, 
                aes(y = NULL, ymin = lower, ymax = upper, color = Group, fill = Group),
                alpha = .15) 

###################
# Look at model 2:

summary(m2)
confint(m2, parm = "He")
confint(m2, parm = "velocity")

# Plot random effects
# Does not work because the functions doesnt know how to deal with the multiple random slopes
# eff_plot <- broom.mixed::tidy(m2, effects = "ran_vals", conf.int = TRUE)
# eff_plot <- eff_plot %>% dplyr::filter(group == "Group:Param" & term == "He")
# tmp <- lapply(eff_plot$level, function(x) {
#     strsplit(x,":")[[1]]
# })
# eff_plot$Group <- sapply(tmp, function(x) x[1])
# eff_plot$Param <- sapply(tmp, function(x) x[2])
# 
# ggplot(eff_plot, aes(x=estimate, y=Param, colour = Group))+
#     geom_point()+
#     geom_pointrange(aes(xmin = conf.low, xmax = conf.high))+
#     facet_wrap(term~Group)+
#     xlab("Random effect")+
#     geom_vline(xintercept = 0, linetype = "dashed")

# model1 <- MCMCglmm(lag ~ He + Method + ID.area,
#                    random=~us(He):Group:Param, data=mydatatogo, verbose=FALSE,
#                    nitt=1300, burnin=300, thin=1)
# 
# summary(model1)




```


