---
title: "Adaptive potential\nMerge genetic data"
author: "Bioshifts"
output:
    html_document:
        toc: true
        toc_float: true
        number_sections: true
---

# Load libraries and source code

```{r libs}

library(ggplot2)
require(data.table)
library(dplyr)
library(tidyr)
require(parallel)
library(bdc)
library(taxadb)
library(traitdataform)
library(pbapply)

source(here::here("R/clean_taxa_functions.r"))
```

# Load bishifts

```{r bioshifts}
bs_v1 <- fread(here::here('Data/biov1_fixednames.csv'))
bs_v1$spp <- gsub("_"," ",bs_v1$sp_name_std_v1)

bs_v2 <- fread(here::here('Data/biov2_fixednames.csv'))
bs_v2$spp <- gsub("_"," ",bs_v2$sp_name_std_v2)
```

# Load genetic data

```{r gen_data}
mt <- fread(here::here('Data/mtdna.csv'))
ms <- fread(here::here('Data/msat.csv'))
gen_d <- fread(here::here('Data/Deposited_data_genetic_diversity_dekort2021.csv'))
gen_d$spp <-  gsub("_"," ",gen_d$Species)
gen_lf <- fread(here::here('Data/MacroPopGen_Database_final_areas_Lawrence_Fraser_2020.csv'))
gen_lf$spp <-  gsub("_"," ",gen_lf$G_s)

```

## Stats

```{r}

gen_sps <- unique(c(mt$spp, ms$spp, gen_d$spp, gen_lf$spp))

bsi_v1 <- intersect(bs_v1$spp, gen_sps)
bsi_v2 <- intersect(bs_v2$spp, gen_sps)

bsi_tot <- union(bsi_v1,bsi_v2)

# Break down

tot1 <- data.frame(N_species_with_gen_data = length(gen_sps),
                   v1_matches = length(bsi_v1),
                   v2_matches = length(bsi_v2),
                   v3_matches = length(bsi_tot))
tot1
```

There are `r length(gen_sps)` species with gen data before fixing sci names.

`r length(bsi_v1)` species match with bioshifts v1 before fixing sci names.

`r length(bsi_v2)` species match with bioshifts v2 before fixing sci names.

`r length(bsi_tot)` species match before fixing sci names.

# Fix names at the genetic data

```{r}
# Species to find

mycols <- c("species","scientificName","kingdom","phylum","class","order","family","db_code")
splist <- data.frame(matrix(ncol = length(mycols), nrow = length(gen_sps)))
names(splist) <- mycols

splist$reported_name_fixed <- gen_sps
tofind_ <- splist[which(is.na(splist$scientificName)),]
tofind_ <- unique(tofind_$reported_name_fixed)

tofind <- data.frame(matrix(nrow = length(tofind_), ncol = 8))
names(tofind) = c("scientificName", "kingdom", "phylum", "class", "order", "family", "db", "db_code")

tofind <- data.frame(species = tofind_, tofind)

tofind <- tofind %>%
    mutate(across(everything(), as.character))



## GBIF
togo <- tofind[which(is.na(tofind$scientificName)),]

mycols <- c("speciesKey", "kingdom","phylum","class","order","family", "species")

# retrieve sp names
cl <- makeCluster(detectCores()-2)
clusterExport(cl, c("togo","standardize_taxa"))

gbif_names <- pblapply(togo$species, function(x){
    try(standardize_taxa(data.frame(verbatimScientificName = x), 
                         fuzzy = FALSE,
                         silent = TRUE))
}, cl = cl)

stopCluster(cl)

rem <- sapply(gbif_names, class)
if(any(rem=="try-error")){
    rem <- which(rem=="try-error")
    gbif_names <- gbif_names[-rem]
}
gbif_names <- rbindlist(gbif_names)

if(any(is.na(gbif_names$scientificName))){
    gbif_names <- gbif_names[-which(is.na(gbif_names$scientificName)),]
}
gbif_names <- gbif_names[which(gbif_names$taxonRank=='species'),]

# remove duplicates
if(any(duplicated(gbif_names$verbatimScientificName))){
    gbif_names <- gbif_names[-which(duplicated(gbif_names$verbatimScientificName)),]
}

any(!gbif_names$original_search %in% splist$species)

cat("--- Summary ---\n",
    "N taxa:",nrow(togo),"\n",
    "N taxa found:",nrow(gbif_names), "\n",
    "N taxa not found:", nrow(togo)-nrow(gbif_names))

gbif_names <- gbif_names[,c("verbatimScientificName","scientificName","kingdom","phylum","class","order","family","taxonID")]
names(gbif_names) <- c("species","scientificName","kingdom","phylum","class","order","family","db_code")
gbif_names$db <- "gbif"
gbif_names$db_code <- gsub("http://www.gbif.org/species/","",gbif_names$db_code)
gbif_names$db_code <- paste("GBIF:",gbif_names$db_code,sep = "")

# Feed
tofind <- tofind %>% 
    rows_patch(gbif_names, 
               by = "species")

gc()

## Add found species names to the splist


all(tofind$species %in% splist$reported_name_fixed)

if(any(is.na(tofind$scientificName))){
    found <- tofind[-which(is.na(tofind$scientificName)),]
} else {
    found = tofind
}


for(i in 1:length(found$species)){
    tofill <- unique(which(splist$reported_name_fixed == found$species[i]))
    splist$scientificName[tofill] <- found$scientificName[i]
    splist$kingdom[tofill] <- found$kingdom[i]
    splist$phylum[tofill] <-found$phylum[i]
    splist$class[tofill] <- found$class[i]
    splist$order[tofill] <- found$order[i]
    splist$family[tofill] <- found$family[i]
    splist$db[tofill] <- found$db[i]
    splist$db_code[tofill] <- found$db_code[i]
}

any(!splist$scientificName == splist$reported_name_fixed)

splist$spp = splist$reported_name_fixed

### Fix names

any(!mt$spp %in% gen_sps)
any(!mt$spp %in% splist$spp)

any(!ms$spp %in% gen_sps)
any(!ms$spp %in% splist$spp)

any(!gen_d$spp %in% gen_sps)
any(!gen_d$spp %in% splist$spp)

any(!gen_lf$spp %in% gen_sps)
any(!gen_lf$spp %in% splist$spp)


for(i in 1:nrow(mt)){
    mt$spp_new[i] <- splist$scientificName[which(splist$spp == mt$spp[i])]
}
for(i in 1:nrow(ms)){
    ms$spp_new[i] <- splist$scientificName[which(splist$spp == ms$spp[i])]
}
for(i in 1:nrow(gen_d)){
    gen_d$spp_new[i] <- splist$scientificName[which(splist$spp == gen_d$spp[i])]
}
for(i in 1:nrow(gen_lf)){
    gen_lf$spp_new[i] <- splist$scientificName[which(splist$spp == gen_lf$spp[i])]
}
```

## Stats

```{r}
gen_sps <- unique(c(mt$spp_new, ms$spp_new, gen_d$spp_new, gen_lf$spp_new))

bsi_v1 <- intersect(bs_v1$spp, gen_sps)
bsi_v2 <- intersect(bs_v2$spp, gen_sps)

bsi_tot <- union(bsi_v1,bsi_v2)

length(gen_sps) # 1821 spp with gen data after fixing sci names

length(bsi_v1) # 345 species match with bioshifts v1 after fixing sci names
length(bsi_v2) # 351 species match with bioshifts v2 after fixing sci names
length(bsi_tot) # 395 species match after fixing sci names

# Break down

tot2 <- data.frame(N_species_with_gen_data = length(gen_sps),
                   v1_matches = length(bsi_v1),
                   v2_matches = length(bsi_v2),
                   v3_matches = length(bsi_tot))

total <- data.frame(cbind(t(tot1),t(tot2)))
names(total) <- c("Before_tax_harmonization","After_tax_harmonization")
total
```

# Which species

```{r}

toplot = data.frame(table(splist$class))

ggplot(toplot, aes(x = Var1, y = Freq))+
    ggtitle("N species by Class")+
    geom_bar(stat="identity")+
    geom_text(aes(y=Freq, label=Freq), 
                  hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

toplot <- splist %>%
    group_by(family) %>%
    summarise(N = length(family))
level_order <- toplot$family[order(toplot$N)]

ggplot(toplot, aes(x=factor(family, level = level_order), y = N))+
    ggtitle("N species by Family")+
    geom_col()+
    theme_classic()+
    coord_flip()+
    xlab("")
```

# Explore relatioships

## *Malin's data*

```{r}

bs_v2$SHIFT <- as.numeric(bs_v2$`Conversion to km/dec`)

bs_v1$sp_name_std_v1 <- gsub("_"," ",bs_v1$sp_name_std_v1)
bs_v2$sp_name_std_v2 <- gsub("_"," ",bs_v2$sp_name_std_v2)

bs_v1_ms <- merge(bs_v1, ms, by.x = "sp_name_std_v1", by.y = "spp_new")
bs_v1_mt <- merge(bs_v1, mt, by.x = "sp_name_std_v1", by.y = "spp_new")

bs_v2_ms <- merge(bs_v2, ms, by.x = "sp_name_std_v2", by.y = "spp_new", allow.cartesian=TRUE)
bs_v2_mt <- merge(bs_v2, mt, by.x = "sp_name_std_v2", by.y = "spp_new")

ggplot(bs_v1_ms, aes(x = He, y = SHIFT))+
    ggtitle("Malin's data (Micro satelite (He))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "gam")

ggplot(bs_v1_ms, aes(x = Hese, y = SHIFT))+
    ggtitle("Malin's data (Micro satelite  (Hese))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "gam")

ggplot(bs_v1_mt, aes(x = He, y = SHIFT))+
    ggtitle("Malin's data (Mitocondrial DNA (He))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "gam")

ggplot(bs_v1_mt, aes(x = Hese, y = SHIFT))+
    ggtitle("Malin's data (Mitocondrial DNA (Hese))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "gam")

ggplot(bs_v2_ms, aes(x = He, y = SHIFT))+
    ggtitle("Malin's data (Micro satelite (He))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "gam")

ggplot(bs_v2_ms, aes(x = Hese, y = SHIFT))+
    ggtitle("Malin's data (Micro satelite (Hese))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "gam")

ggplot(bs_v2_mt, aes(x = He, y = SHIFT))+
    ggtitle("Malin's data (Mitocondrial DNA (He))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "gam")

ggplot(bs_v2_mt, aes(x = Hese, y = SHIFT))+
    ggtitle("Malin's data (Mitocondrial DNA (Hese))\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "gam")


```

## *De Kort 2021 data*

```{r}
bs_v1_d <- merge(bs_v1[,c("sp_name_std_v1","SHIFT", "Class")], 
                 gen_d[,c("GDp","spp_new")], 
                 by.x = "sp_name_std_v1", by.y = "spp_new")

bs_v2_d <- merge(bs_v2[,c("sp_name_std_v2","SHIFT", "class")], 
                 gen_d[,c("GDp","spp_new")], 
                 by.x = "sp_name_std_v2", by.y = "spp_new", 
                 allow.cartesian=TRUE)

ggplot(na.omit(bs_v1_d) %>%
           filter(!SHIFT < -40), 
       aes(x = GDp, y = SHIFT))+
    ggtitle("De Kort 2021 (Genetic diversity)\nBioshifts v1")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "lm") +
    facet_wrap(.~Class, scales = "free")

ggplot(na.omit(bs_v2_d), aes(x = GDp, y = SHIFT))+
    ggtitle("De Kort 2021 (Genetic diversity)\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "lm") +
    facet_wrap(.~class, scales = "free")
```

## *De Lawrence & Fraser 2020 data*

```{r}
bs_v1_lf <- merge(bs_v1[,c("sp_name_std_v1","SHIFT", "Class")], 
                  gen_lf[,c("FST","spp_new")], 
                  by.x = "sp_name_std_v1", by.y = "spp_new")
bs_v1_lf$FST <- as.numeric(bs_v1_lf$FST)

bs_v2_lf <- merge(bs_v2[,c("sp_name_std_v2","SHIFT", "class")],
                  gen_lf[,c("FST","spp_new")], 
                  by.x = "sp_name_std_v2", by.y = "spp_new", allow.cartesian=TRUE)
bs_v2_lf$FST <- as.numeric(bs_v2_lf$FST)

ggplot(bs_v1_lf, aes(x = FST, y = SHIFT))+
    ggtitle("Lawrence & Fraser 2020 (FST)\nBioshifts v1")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = Class), method = "lm") +
    facet_wrap(.~Class, scales = "free")

ggplot(bs_v2_lf, aes(x = FST, y = SHIFT))+
    ggtitle("Lawrence & Fraser 2020 (FST)\nBioshifts v2")+
    geom_point(alpha = .3)+
    theme_bw()+
    geom_smooth(aes(color = class), method = "lm") +
    facet_wrap(.~class, scales = "free")
```

**END**
