---
title: "Adaptive potential\nMerge genetic data"
author: "Bioshifts group"
format: 
  html:
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Code"
---

# Load libraries and source code

```{r libs, echo=FALSE}
rm(list=ls())
gc()

library(ggplot2)
require(data.table)
library(dplyr)
library(tidyr)
require(parallel)
library(bdc)
library(taxadb)
library(traitdataform)
library(pbapply)
library(tidyverse)
library(readxl)
library(lme4)
library(coefplot)
library(sjPlot)
library(sjmisc)
library(effects)
library(rgdal)
library(pbapply)
library(maptools)
library(rgeos)
library(terra)
library(nlme)

```

```{r func}
source(here::here("R/clean_taxa_functions.r"))
```

# Load species list

```{r splist}

splist <- read.csv(here::here("Data/splist.csv"), header = T)
# remove duplicated sp_names
splist <- splist %>%
    dplyr::select(scientificName,kingdom,phylum,class,order,family,db) %>%
    filter(!duplicated(scientificName))

```

# Load bioshifts

## Load v1

```{r v1}
biov1 <- read.csv(here::here("Data/biov1_fixednames.csv"), header = T)
# Fix references in biov1
biov1$Article_ID <- gsub("[^0-9.-]", "", biov1$Article_ID)
biov1$Article_ID <- as.numeric(biov1$Article_ID)
biov1$sp_name_std_v1 <- gsub("_"," ",biov1$sp_name_std_v1)
biov1 <- biov1 %>%
    dplyr::select(ID,Article_ID,Study_ID,Type,Param,Trend,SHIFT,UNIT,DUR,v.lat.mean,v.ele.mean,START,DUR,
                  START,END,group,ECO,Sampling,Uncertainty_Distribution,Uncertainty_Parameter,
                  Grain_size,Data,ID.area,
                  Phylum,Class,Order,Family,Genus,sp_name_std_v1) %>% # select columns
    mutate(
        Type = case_when(
            Type=="HOR" ~ "LAT",
            TRUE ~ as.character(Type)),
        Data = case_when(
            Data=="occurence-based" ~ "occurrence-based",
            TRUE ~ as.character(Data)),
        Species = sp_name_std_v1) %>% 
    mutate( # from m/year to km/year
        kmyr = case_when(
            UNIT == "m/year" ~ SHIFT/1000, 
            UNIT == "km/year" ~ SHIFT, 
            TRUE ~ NA_real_))

biov1 <- biov1 %>%
    filter(!is.na(sp_name_std_v1))

biov1$spp <- biov1$sp_name_std_v1

all(biov1$sp_name_std_v1 %in% splist$scientificName)

```

## Classify species

```{r classify}

# Class species as Terrestrial, Marine or Freshwater
Terv1 <- unique(biov1$sp_name_std_v1[which(biov1$ECO == "T")])
Terrestrials = unique(c(Terv1))

Mar1 <- unique(biov1$sp_name_std_v1[which(biov1$ECO == "M")])
Marine = unique(c(Mar1))

# Freshwater fish
FFishv1 <- unique(biov1$sp_name_std_v1[(biov1$Class == "Actinopterygii" | biov1$Class == "Cephalaspidomorphi") & biov1$ECO == "T"])
FFish = unique(c(FFishv1))
# Marine fish
MFishv1 <- unique(biov1$sp_name_std_v1[biov1$Class == "Actinopterygii" & biov1$ECO == "M"])
MFish = unique(c(MFishv1))

splist$ECO = NA
splist$ECO[which(splist$scientificName %in% Terrestrials)] <- "T"
splist$ECO[which(splist$scientificName %in% Marine)] <- "M"
splist$ECO[which(splist$scientificName %in% MFish)] <- "M"

biov1$ECO = NA
biov1$ECO[which(biov1$sp_name_std_v1 %in% Terrestrials)] <- "T"
biov1$ECO[which(biov1$sp_name_std_v1 %in% Marine)] <- "M"
biov1$ECO[which(biov1$sp_name_std_v1 %in% MFish)] <- "M"

splist$Group = NA
splist$Group[which(splist$class == "Phaeophyceae")] <- "Chromista"
splist$kingdom[which(splist$class == "Phaeophyceae")] <- "Chromista"
splist$Group[which(splist$phylum == "Rhodophyta")] <- "Seaweed"
splist$kingdom[which(splist$phylum == "Rhodophyta")] <- "Plantae"
splist$Group[which(splist$family == "Elminiidae")] <- "Barnacles"
splist$Group[which(splist$kingdom == "Bacteria")] <- "Bacteria"
splist$Group[which(splist$class == "Holothuroidea")] <- "Sea cucumber"
splist$Group[which(splist$class == "Aves")] <- "Bird"
splist$Group[which(splist$class == "Insecta")] <- "Insect"
splist$Group[which(splist$class == "Mammalia")] <- "Mammal"
splist$Group[which(splist$class == "Arachnida")] <- "Spider"
splist$kingdom[which(splist$kingdom == "Viridiplantae")] <- "Plantae"
splist$kingdom[which(splist$phylum == "Tracheophyta")] <- "Plantae"
splist$Group[which(splist$kingdom == "Plantae")] <- "Plant"
splist$Group[which(splist$class == "Hydrozoa")] <- "Hydrozoa"
splist$Group[which(splist$class == "Anthozoa")] <- "Sea anemones and corals"
splist$Group[which(splist$class == "Polychaeta")] <- "Polychaetes"
splist$Group[which(splist$phylum == "Mollusca")] <- "Molluscs"
splist$Group[which(splist$class == "Malacostraca")] <- "Crustacean"
splist$Group[which(splist$class == "Hexanauplia")] <- "Crustacean"
splist$Group[which(splist$class == "Maxillopoda")] <- "Crustacean"
splist$Group[which(splist$class == "Ostracoda")] <- "Crustacean"
splist$Group[which(splist$class == "Branchiopoda")] <- "Crustacean"
splist$Group[which(splist$class == "Asteroidea")] <- "Starfish"
splist$Group[which(splist$class == "Ascidiacea")] <- "Ascidians tunicates and sea squirts"
splist$class[which(splist$class == "Actinopteri")] <- "Actinopterygii"
splist$Group[which(splist$class == "Actinopterygii")] <- "Fish"
splist$Group[which(splist$class == "Elasmobranchii")] <- "Fish"
splist$Group[which(splist$order == "Perciformes")] <- "Fish"
splist$Group[which(splist$class == "Chondrichthyes")] <- "Fish"
splist$Group[which(splist$class == "Holocephali")] <- "Fish"
splist$Group[which(splist$class == "Cephalaspidomorphi")] <- "Fish"
splist$Group[which(splist$class == "Echinoidea")] <- "Sea urchin"
splist$Group[which(splist$class == "Crinoidea")] <- "Crinoid"
splist$Group[which(splist$class == "Holothuroidea")] <- "Sea cucumber"
splist$Group[which(splist$class == "Reptilia")] <- "Reptile"
splist$Group[which(splist$order == "Squamata")] <- "Reptile"
splist$Group[which(splist$class == "Ophiuroidea")] <- "Brittle stars"
splist$Group[which(splist$class == "Chilopoda")] <- "Centipedes"
splist$Group[which(splist$class == "Diplopoda")] <- "Millipedes"
splist$Group[which(splist$class == "Amphibia")] <- "Amphibian"
splist$Group[which(splist$kingdom == "Fungi")] <- "Fungi"
splist$Group[which(splist$order == "Balanomorpha")] <- "Barnacles"
splist$Group[which(splist$phylum == "Nematoda")] <- "Nematodes"
splist$Group[which(splist$class == "Myxini")] <- "Hagfish"
splist$Group[which(splist$kingdom == "Chromista")] <- "Chromista"

######################################
biov1 <- merge(biov1[,-which(names(biov1) %in% c("Group","ECO"))], 
               splist[,c("Group","ECO","scientificName")], 
               by.x = "sp_name_std_v1", by.y = "scientificName", 
               all.x = T)

```

## Filter LAT / ELE shifts

```{r filter}

# v1
biov1 <- biov1 %>%
    filter(Type %in% c("ELE","LAT"))  # Use LAT ELE shifts

# splist
sps <- unique(biov1$sp_name)
splist <- splist %>% filter(scientificName %in% sps)

```

## Filter species identified to the genus level or cf.

```{r clean}

if(any(grep("sp[.]",biov1$sp_reported_name_v1))){
    biov1 <- biov1 %>% filter(!grepl("sp[.]",sp_reported_name_v1))
}
if(any(grep("sp[.]",biov1$sp_name_std_v1))){
    biov1 <- biov1 %>% filter(!grepl("sp[.]",sp_name_std_v1))
}
if(any(grep("cf[.]",biov1$sp_reported_name_v1))){
    biov1 <- biov1 %>% filter(!grepl("cf[.]",sp_reported_name_v1))
}
if(any(grep("cf[.]",biov1$sp_name_std_v1))){
    biov1 <- biov1 %>% filter(!grepl("cf[.]",sp_name_std_v1))
}

#remove freshwater fishes
biov1 <- biov1[-which(biov1$Class == "Actinopterygii" & biov1$ECO=="T"),]

#remove marine birds
biov1 <- biov1[-which(biov1$Class == "Aves" & biov1$ECO=="M"),]


if(any(grep("sp[.]",splist$scientificName))){
    splist <- splist %>% filter(!grepl("sp[.]",scientificName))
}
if(any(grep("sp[.]",splist$scientificName))){
    splist <- splist %>% filter(!grepl("sp[.]",scientificName))
}
if(any(grep("cf[.]",splist$scientificName))){
    splist <- splist %>% filter(!grepl("cf[.]",scientificName))
}
if(any(grep("cf[.]",splist$scientificName))){
    splist <- splist %>% filter(!grepl("cf[.]",scientificName))
}


```

## Reorganize methodological variables
```{r v1}

unique(biov1$Data)
biov1$Data[biov1$Data=="occurence-based"] <- "occurrence-based"
biov1$Data <- factor(biov1$Data, levels = unique(biov1$Data))
table(biov1$Data)

unique(biov1$Sampling)
biov1$Sampling = ifelse(biov1$Sampling %in% c("IRR","MULTIPLE(continuous)"),"MULTIPLE", biov1$Sampling)
biov1$Sampling <- ordered(biov1$Sampling, 
                          levels = c("TWO","MULTIPLE","CONTINUOUS"))
table(biov1$Sampling)

unique(biov1$Grain_size)
biov1$Grain_size <- ifelse(biov1$Grain_size %in% c("large","very_large"),"large",biov1$Grain_size)
biov1$Grain_size <- ordered(biov1$Grain_size, 
                            levels = c("small","moderate","large"))
table(biov1$Grain_size)

unique(biov1$Uncertainty_Distribution)
biov1$Uncertainty_Distribution <- ifelse(biov1$Uncertainty_Distribution %in% c("RESAMPLING","RESAMPLING(same)"),"RESAMPLING",
                                         ifelse(biov1$Uncertainty_Distribution %in% c("MODEL","MODEL+RESAMPLING(same)","RESAMPLING+MODEL"),"MODEL",
                                                ifelse(biov1$Uncertainty_Distribution %in% c("DETECTABILITY","RESAMPLING(same)+DETECTABILITY"),"DETECTABILITY",
                                                       biov1$Uncertainty_Distribution)))
biov1$Uncertainty_Distribution <- ifelse(biov1$Uncertainty_Distribution == "RAW","OPPORTUNISTIC","PROCESSED")  
table(biov1$Uncertainty_Distribution)

#transform study area
biov1$ID.area <- log(biov1$ID.area)

```

## Direction of shift
```{r v1}
biov1 <- biov1 %>%
    mutate(shift_sign = ifelse(SHIFT>0,"pos","neg"),
           velocity = ifelse(Type == "LAT", v.lat.mean, v.ele.mean),
           vel_sign = ifelse(velocity>0,"pos","neg"),
           shift_vel_sign = paste0(shift_sign,vel_sign))

ggplot(biov1, aes(x=shift_vel_sign))+
    geom_bar()+
    coord_flip()+
    facet_wrap(Type~Param)

```

## Get centroids of study areas
This will be used to merge genetic data with bioshifts based on distance of observations
```{r v1}
# # The input file geodatabase
# fgdb <- "C:/Users/brunn/NextCloud/Bioshifts/Study_Areas.gdb"
# 
# # List all feature classes in a file geodatabase
# fc_list <- ogrListLayers(fgdb)
# 
# # Get centroids
# cl <- makeCluster(detectCores()-1)
# clusterExport(cl, c("readOGR", "gCentroid", "fgdb"))
# 
# centroids <- pblapply(fc_list, function(x){
#     fc <- readOGR(dsn=fgdb,layer=x)
#     c <- gCentroid(fc)
#     data.frame(c)
# }, cl = cl)
# 
# stopCluster(cl)
# 
# centroids <- do.call("rbind",centroids)
# centroids$ID <- fc_list
# 
# write.csv(centroids, "Data/centroids_study_areas.csv", row.names = FALSE)

centroids <- read.csv(here::here("Data/centroids_study_areas.csv"))
names(centroids) <- c("lat","long","ID")

biov1 <- merge(biov1, centroids, by = "ID")

```

# Calculated residual from shifts
## Full model
```{r shift_res_full, fig.width=5, fig.height=5}

# biov1 %>%
#     group_by(Grain_size) %>%
#     summarise(mean(ID.area))
# # One observation in bioshifts has no grain data
# 
# #transforming continuous method variables in qualitative variables => two benefits: (1) all variables area qualitative (2) better than quantitative in case the effect is not linear
# q1=quantile(biov1$START,probs=c(0,0.25,0.5,0.75,1))
# biov1$StartF=cut(biov1$START,breaks=q1,include.lowest=T)
# 
# q1=quantile(biov1$ID.area,probs=c(0,0.25,0.5,0.75,1))
# biov1$ID.areaF=cut(biov1$ID.area,breaks=q1,include.lowest=T)
# 
# q1=quantile(biov1$DUR,probs=c(0,0.25,0.5,0.75,1))
# biov1$DURF=cut(biov1$DUR,breaks=q1,include.lowest=T)
# 
# biov1$Grain_sizeF <- as.numeric(biov1$Grain_size)
# 
# #MODELS with methods only
# meth <- biov1[,names(biov1) %in% c("Param","Data","Sampling","Grain_size","Uncertainty_Distribution","Uncertainty_Parameter","START","DUR","ID.area")] #remove "n" as can capture diversity gradients
# 
# meth <- na.omit(meth)
# 
# biov1$Grain_size <- as.numeric(biov1$Grain_size)
# 
# # convert factors to numeric
# meth[,names(meth) %in% c("ID","ID.areaF","DURF","StartF","Param","Data","Sampling","Uncertainty_Distribution","Uncertainty_Parameter","Grain_size")] <- apply(meth[,names(meth) %in% c("ID","ID.areaF","DURF","StartF","Param","Data","Sampling","Uncertainty_Distribution","Uncertainty_Parameter","Grain_size")],2,function(x) as.numeric(factor(x)))
# cormat <- cor(meth)
# corrplot::corrplot(cormat)
# #id.area and grain size are correlated so pick only area
# #start and duration are correlated so pick only dur
# # Or, we can use categorical version of variables
# 
# # My models
# # categorical version of continual variables + random effects (ID + taxa)
# # issue1: do we want to account for taxonomy?
# # issue2: Is ID redundant with methodological variables?
# model1 <- "SHIFT ~ Param + DURF + StartF + ID.areaF + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution + (1|spp) + (1|ID)"
# # continuous variables + random effects (ID + taxa)
# # issue1: do we want to account for taxonomy?
# # issue2: Is ID redundant with methodological variables?
# model2 <- "SHIFT ~ Param + DUR + START + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution + (1|spp) + (1|ID)"
# # continuous variables + random effects (ID + taxa)
# # issue1: do we want to account for taxonomy?
# # issue2: Is ID redundant with methodological variables?
# model3 <- "SHIFT ~ DUR + START + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution"
# 
# model3 <- "SHIFT ~ Param + (1|ID) + (1|spp)"
# model4 <- "SHIFT ~ DUR + START + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution"
# 
# # Latitude
# my_vars <- c("spp","SHIFT","velocity","Class","Family","Genus","Group","shift_vel_sign","lat","long",
#              "ID","Type","Param","Data","Sampling","Grain_size","Uncertainty_Distribution","Uncertainty_Parameter","START","DUR","ID.area","DURF","StartF","ID.areaF")
# 
# data.lat <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "LAT") %>%
#     na.omit
# 
# # add random variation to longlat
# noise = rnorm(nrow(data.lat),0,0.0001)
# data.lat$lat = data.lat$lat+noise
# data.lat$long = data.lat$long+noise
# 
# mod.meth.lat3 <- lmer(model3,
#                      REML=TRUE,na.action="na.fail",
#                      control = lmerControl(optimizer ='optimx', 
#                                            optCtrl=list(method='nlminb')),
#                      data = data.lat)
# MuMIn::r.squaredGLMM(mod.meth.lat3)
# # m1
# # R2m       R2c
# # 0.00451421 0.8754512
# 
# # m2
# # R2m       R2c
# # 0.001147401 0.873946
# 
# # m3
# # R2m       R2c
# # 0.1844739 0.9681493
# 
# AIC(mod.meth.lat)
# 
# mod.meth.lat <- gls(as.formula(model4),
#                     correlation = corSpatial(form = ~ long+lat),
#                     method = "REML",na.action="na.fail",
#                     control = glsControl(opt  = 'nlminb'),
#                     data = data.lat)
# MuMIn::r.squaredGLMM(mod.meth.lat)
# 
# # summary(mod.meth.lat)
# 
# # inspect residuals
# plot(mod.meth.lat)
# 
# predicted.shift <- residuals(mod.meth.lat)
# 
# nrow(data.lat)
# length(predicted.shift)
# 
# data.lat$shift_corrected <- predicted.shift
# 
# # Elevation
# 
# # shift
# data.ele <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "ELE") %>%
#     na.omit
# 
# mod.meth.ele <- lmer(model3,
#                      REML=TRUE,na.action="na.fail",
#                      control = lmerControl(optimizer ='optimx', 
#                                            optCtrl=list(method='nlminb')),
#                      data = data.ele)
# MuMIn::r.squaredGLMM(mod.meth.ele)
# # m1
# # R2m       R2c
# # 0.02026562 0.1396005
# 
# # m2
# # R2m       R2c
# # 0.009871563 0.1304724
# 
# # m3
# # R2m       R2c
# # 0.06902197 0.1877013
# 
# AIC(mod.meth.ele)
# 
# # summary(mod.meth.lat)
# 
# # inspect residuals
# plot(mod.meth.ele)
# 
# predicted.shift <- residuals(mod.meth.ele)
# 
# nrow(data.ele)
# length(predicted.shift)
# 
# data.ele$shift_corrected <- predicted.shift
# 
# # Merge
# biov1_new <- rbind(data.lat,data.ele)

```
## Correlation between raw and corrected shifts

```{r cor_model_full, fig.width=10, fig.height=3}

# {
#     par(mfrow = c(1,3))
#     plot(shift_corrected~SHIFT, 
#          biov1_new %>% filter(Type == "LAT"),
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          main = "Latitude data")
#     plot(shift_corrected~SHIFT, 
#          biov1_new %>% filter(Type == "ELE"), 
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          main = "Elevation data")
#     plot(shift_corrected~SHIFT, 
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          biov1_new, 
#          main = "Full data")
# }
```

Inspecting the residuals of the full models vs the subset models, it seems the residuals tend of the 

## Subset models
```{r shift_res_subset, fig.width=5, fig.height=5}

# # Latitude
# # LE
# data.lat.le <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "LAT") %>%
#     filter(Param == "LE") %>%
#     na.omit
# 
# mod.meth.lat.le <- lmer(model3,
#                         REML=TRUE,na.action="na.fail",
#                         control = lmerControl(optimizer ='optimx', 
#                                               optCtrl=list(method='nlminb')),
#                         data = data.lat.le)
# MuMIn::r.squaredGLMM(mod.meth.lat.le)
# # R2m       R2c
# # 0.2191011 0.3706496
# 
# AIC(mod.meth.lat.le)
# 
# # summary(mod.meth.lat.le)
# 
# # inspect residuals
# plot(mod.meth.lat.le)
# 
# predicted.shift <- residuals(mod.meth.lat.le)
# 
# nrow(data.lat.le)
# length(predicted.shift)
# 
# data.lat.le$shift_corrected <- predicted.shift
# 
# # TE
# data.lat.te <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "LAT") %>%
#     filter(Param == "TE") %>%
#     na.omit
# 
# mod.meth.lat.te <- lmer(model3,
#                         REML=TRUE,na.action="na.fail",
#                         control = lmerControl(optimizer ='optimx', 
#                                               optCtrl=list(method='nlminb')),
#                         data = data.lat.te)
# MuMIn::r.squaredGLMM(mod.meth.lat.te)
# # R2m       R2c
# # 0.3411776 0.9877747
# 
# AIC(mod.meth.lat.te)
# 
# # summary(mod.meth.lat.te)
# 
# # inspect residuals
# plot(mod.meth.lat.te)
# 
# predicted.shift <- residuals(mod.meth.lat.te)
# 
# nrow(data.lat.te)
# length(predicted.shift)
# 
# data.lat.te$shift_corrected <- predicted.shift
# 
# # O
# data.lat.o <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "LAT") %>%
#     filter(Param == "O") %>%
#     na.omit
# 
# mod.meth.lat.o <- lmer(model3,
#                        REML=TRUE,na.action="na.fail",
#                        control = lmerControl(optimizer ='optimx', 
#                                              optCtrl=list(method='nlminb')),
#                        data = data.lat.o)
# MuMIn::r.squaredGLMM(mod.meth.lat.o)
# # R2m       R2c
# # 0.2745957 0.9725265
# 
# AIC(mod.meth.lat.o)
# 
# # summary(mod.meth.lat.o)
# 
# # inspect residuals
# plot(mod.meth.lat.o)
# 
# predicod.shift <- residuals(mod.meth.lat.o)
# 
# nrow(data.lat.o)
# length(predicod.shift)
# 
# data.lat.o$shift_corrected <- predicod.shift
# 
# # Elevation
# 
# # LE
# data.ele.le <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "ELE") %>%
#     filter(Param == "LE") %>%
#     na.omit
# 
# mod.meth.ele.le <- lmer(model3,
#                         REML=TRUE,na.action="na.fail",
#                         control = lmerControl(optimizer ='optimx', 
#                                               optCtrl=list(method='nlminb')),
#                         data = data.ele.le)
# MuMIn::r.squaredGLMM(mod.meth.ele.le)
# # R2m       R2c
# # 0.05451751 0.0724277
# 
# AIC(mod.meth.ele.le)
# 
# # summary(mod.meth.ele.le)
# 
# # inspect residuals
# plot(mod.meth.ele.le)
# 
# predicted.shift <- residuals(mod.meth.ele.le)
# 
# nrow(data.ele.le)
# length(predicted.shift)
# 
# data.ele.le$shift_corrected <- predicted.shift
# 
# # TE
# data.ele.te <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "ELE") %>%
#     filter(Param == "TE") %>%
#     na.omit
# 
# mod.meth.ele.te <- lmer(model3,
#                         REML=TRUE,na.action="na.fail",
#                         control = lmerControl(optimizer ='optimx', 
#                                               optCtrl=list(method='nlminb')),
#                         data = data.ele.te)
# MuMIn::r.squaredGLMM(mod.meth.ele.te)
# # R2m       R2c
# # 0.07846136 0.597949
# 
# AIC(mod.meth.ele.te)
# 
# # summary(mod.meth.ele.te)
# 
# # inspect residuals
# plot(mod.meth.ele.te)
# 
# predicted.shift <- residuals(mod.meth.ele.te)
# 
# nrow(data.ele.te)
# length(predicted.shift)
# 
# data.ele.te$shift_corrected <- predicted.shift
# 
# # O
# data.ele.o <- biov1 %>%
#     select(my_vars) %>%
#     filter(Type == "ELE") %>%
#     filter(Param == "O") %>%
#     na.omit
# 
# mod.meth.ele.o <- lmer(model3,
#                        REML=TRUE,na.action="na.fail",
#                        control = lmerControl(optimizer ='optimx', 
#                                              optCtrl=list(method='nlminb')),
#                        data = data.ele.o)
# MuMIn::r.squaredGLMM(mod.meth.ele.o)
# # R2m       R2c
# # 0.05301482 0.1656916
# 
# AIC(mod.meth.ele.o)
# 
# # summary(mod.meth.ele.o)
# 
# # inspect residuals
# plot(mod.meth.ele.o)
# 
# predicod.shift <- residuals(mod.meth.ele.o)
# 
# nrow(data.ele.o)
# length(predicod.shift)
# 
# data.ele.o$shift_corrected <- predicod.shift
# 
# 
# # Merge
# biov1_new2 <- rbind(data.lat.le,
#                     data.lat.o,
#                     data.lat.te,
#                     data.ele.le,
#                     data.ele.o,
#                     data.ele.te)
# 
# # nrow(biov1 %>% filter(shift_sign == "neg"))
# # 
# # 
# # a=c(1:1000)
# # noise=rnorm(length(a), 0.0001, 0.0001)
# # noise=rnorm(length(a), 0, 0.0001)
# # hist(noise)
# # test <- data.frame(a, b=rev(a)+noise, c=a)
# # 
# # plot(test$a,test$b)
# # 
# # m <- lm(a~b,test)
# # plot(m)
# # summary(m)
# # 
# # resm <- residuals(m)
# # hist(resm)
# # 
# # plot(test$a,round(resm,5))
# # 
# # plot(noise, resm)
# # plot(test$a,test$b)
# # lines(test$a, test$a)

```

## Correlation between raw and corrected shifts

```{r cor_model_subset, fig.width=10, fig.height=3}

# {
#     par(mfrow = c(1,3))
#     plot(shift_corrected~SHIFT, 
#          biov1_new2 %>% filter(Type == "LAT"),
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          main = "Latitude data")
#     plot(shift_corrected~SHIFT, 
#          biov1_new2 %>% filter(Type == "ELE"), 
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          main = "Elevation data")
#     plot(shift_corrected~SHIFT, 
#          ylab = "Predicted shift",
#          xlab = "Observed shift",
#          biov1_new2, 
#          main = "Full data")
# }


```
## Select the database
Based on the dispersion of the observed vs residual shifts, both modelling approaches seem similar. I will then select the more simple modelling approach
```{r}
# biov1 = biov1_new
```

# Calculate lags 

```{r lags}

# calculate lags
# positive values are a lag (range shift lower smaller than expected or in opposite directions) and negative values are range shift larger than expected
biov1$lag <- biov1$velocity-biov1$SHIFT
position = which(biov1$vel_sign == "neg")
biov1$lag[position] <- biov1$lag[position] * -1 # Any negative velocity means the sign of lag has to shift.

# Lag 2
# calculate lags based on "corrected" shifts
# biov1$lag2 <- biov1$velocity-biov1$shift_corrected
# position = which(biov1$vel_sign == "neg")
# biov1$lag2[position] <- biov1$lag2[position] * -1 # Any negative velocity means the sign of lag has to shift.

# Lag 3
# When shift is in opposite sign of velocity, lag == velocity
biov1$lag3 = biov1$lag
position <- which(biov1$shift_vel_sign == "posneg" | biov1$shift_vel_sign == "negpos")
biov1$lag3[position] <- biov1$velocity[position]

# Lag 4
# apply lag 3 to lag 2
# biov1$lag4 = biov1$lag2
# position <- which(biov1$shift_vel_sign == "posneg" | biov1$shift_vel_sign == "negpos")
# biov1$lag4[position] <- biov1$velocity[position]


# {
#     par(mfrow=c(2,2))
#     plot(lag~velocity, biov1)
#     plot(lag2~velocity, biov1)
#     plot(lag3~velocity, biov1)
#     plot(lag4~velocity, biov1)
# }
# 
# ggplot() +
#     geom_density(data = biov1, aes(x = lag), fill = "blue", color = "blue", alpha = .2) +
#     geom_density(data = biov1, aes(x = lag2), fill = "red", color = "red", alpha = .2) +
#     geom_density(data = biov1, aes(x = lag3), fill = "green", color = "green", alpha = .2) +
#     geom_density(data = biov1, aes(x = lag4), fill = "yellow", color = "yellow", alpha = .2)

ggplot(biov1 %>%
           filter(shift_vel_sign == 'pospos'), 
       aes(x=abs(lag3), fill = Param, color = Param))+
    geom_histogram(alpha = .5)+
    facet_wrap(.~Type)

```



# Load genetic data

```{r gen_data, eval=FALSE}

# Genetic diversity
mt <- fread(here::here('Data/mtdna.csv'))
mt$spp <- Clean_Names(mt$spp)

ms <- fread(here::here('Data/msat.csv'))
ms$spp <- Clean_Names(ms$spp)

gen_d <- fread(here::here('Data/Deposited_data_genetic_diversity_dekort2021.csv'))
gen_d$spp <-  gsub("_"," ",gen_d$Species)
gen_d$spp <- Clean_Names(gen_d$spp)

gen_lf <- read_excel(here::here("Data/MacroPopGen_Database_final-v0.2.xlsx"), sheet = "MacroPopGen_Database")
gen_lf$spp <-  gsub("_"," ",gen_lf$G_s)
gen_lf$spp <- Clean_Names(gen_lf$spp)

# Phenotypic rates
pheno <- fread(here::here('Data/PROCEEDv6_RatesDB.csv'))
pheno$spp <- Clean_Names(pheno$sp_ncbi)

length(unique(gen_d$spp))
length(unique(gen_lf$spp))
length(unique(mt$spp))
length(unique(ms$spp))
length(unique(pheno$spp))

```

## Stats

```{r stats_gen, eval=FALSE}

# Gen
gen_sps <- unique(c(mt$spp, ms$spp, gen_d$spp, gen_lf$spp))

bsi_v1 <- intersect(biov1$spp, gen_sps)
bsi_v2 <- intersect(biov2$spp, gen_sps)
bsi_v3 <- intersect(biov3$spp, gen_sps)

# Break down
tot1 <- data.frame(N_species_with_gen_data = length(gen_sps),
                   v1_matches = length(bsi_v1),
                   v2_matches = length(bsi_v2),
                   v3_matches = length(bsi_v3))
tot1

# phenotypic
phen_sps <- unique(pheno$spp)

bsj_v1 <- intersect(biov1$spp, phen_sps)
bsj_v2 <- intersect(biov2$spp, phen_sps)
bsj_v3 <- intersect(biov3$spp, phen_sps)

# Break down
tot2 <- data.frame(N_species_with_gen_data = length(phen_sps),
                   v1_matches = length(bsj_v1),
                   v2_matches = length(bsj_v2),
                   v3_matches = length(bsj_v3))
tot2

```


## Fix names at the genetic data

```{r fixnames, eval=FALSE}

# Species to find
mycols <- c("species","scientificName","kingdom","phylum","class","order","family","db_code")
splist <- data.frame(matrix(ncol = length(mycols), nrow = length(unique(c(gen_sps,phen_sps)))))
names(splist) <- mycols

splist$reported_name_fixed <- unique(c(gen_sps,phen_sps))
tofind_ <- splist[which(is.na(splist$scientificName)),]
tofind_ <- unique(tofind_$reported_name_fixed)

tofind <- data.frame(matrix(nrow = length(tofind_), ncol = 8))
names(tofind) = c("scientificName", "kingdom", "phylum", "class", "order", "family", "db", "db_code")

tofind <- data.frame(species = tofind_, tofind)

tofind <- tofind %>%
    mutate(across(everything(), as.character))

# retrieve sp names
sp_names_found <- Find_Sci_Names(spnames = tofind$species)

# Feed
tofind <- tofind %>% 
    rows_patch(sp_names_found, 
               by = "species")

gc()

## Add found species names to the splist
all(tofind$species %in% splist$reported_name_fixed)

if(any(is.na(tofind$scientificName))){
    found <- tofind[-which(is.na(tofind$scientificName)),]
} else {
    found = tofind
}

for(i in 1:length(found$species)){
    tofill <- unique(which(splist$reported_name_fixed == found$species[i]))
    splist$scientificName[tofill] <- found$scientificName[i]
    splist$kingdom[tofill] <- found$kingdom[i]
    splist$phylum[tofill] <-found$phylum[i]
    splist$class[tofill] <- found$class[i]
    splist$order[tofill] <- found$order[i]
    splist$family[tofill] <- found$family[i]
    splist$db[tofill] <- found$db[i]
    splist$db_code[tofill] <- found$db_code[i]
}

splist$spp = splist$reported_name_fixed

# Keep original names for those species we could not find a name at GBIF
pos <- which(is.na(splist$scientificName))
splist$scientificName[pos] <- splist$spp[pos]

any(!mt$spp %in% gen_sps)
any(!mt$spp %in% splist$spp)

any(!ms$spp %in% gen_sps)
any(!ms$spp %in% splist$spp)

any(!gen_d$spp %in% gen_sps)
any(!gen_d$spp %in% splist$spp)

any(!gen_lf$spp %in% gen_sps)
any(!gen_lf$spp %in% splist$spp)

for(i in 1:nrow(mt)){
    mt$spp_new[i] <- splist$scientificName[which(splist$spp == mt$spp[i])]
}
for(i in 1:nrow(ms)){
    ms$spp_new[i] <- splist$scientificName[which(splist$spp == ms$spp[i])]
}
for(i in 1:nrow(gen_d)){
    gen_d$spp_new[i] <- splist$scientificName[which(splist$spp == gen_d$spp[i])]
}
for(i in 1:nrow(gen_lf)){
    gen_lf$spp_new[i] <- splist$scientificName[which(splist$spp == gen_lf$spp[i])]
}
for(i in 1:nrow(gen_lf)){
    pheno$spp_new[i] <- splist$scientificName[which(splist$spp == pheno$spp[i])]
}

write.csv(mt, 
          here::here('Data/mtdna_harmo.csv'), row.names = FALSE)
write.csv(ms, 
          here::here('Data/msat_harmo.csv'), row.names = FALSE)
write.csv(gen_d, 
          here::here('Data/Deposited_data_genetic_diversity_dekort2021_harmo.csv'), row.names = FALSE)
write.csv(gen_lf, 
          here::here('Data/MacroPopGen_Database_final_areas_Lawrence_Fraser_2020_harmo.csv'), row.names = FALSE)
write.csv(pheno, 
          here::here('Data/PROCEEDv6_RatesDB_harmo.csv'), row.names = FALSE)

```

# Load harmonized datasets

```{r load_harmonized_gen}

mt <- fread(here::here('Data/mtdna_harmo.csv'))
ms <- fread(here::here('Data/msat_harmo.csv'))
gen_d <- fread(here::here('Data/Deposited_data_genetic_diversity_dekort2021_harmo.csv'))
gen_d$spp <-  gsub("_"," ",gen_d$Species)
gen_lf <- fread(here::here('Data/MacroPopGen_Database_final_areas_Lawrence_Fraser_2020_harmo.csv'))
gen_lf$spp <-  gsub("_"," ",gen_lf$G_s)
pheno <- fread(here::here('Data/PROCEEDv6_RatesDB_harmo.csv'))
pheno$spp <-  gsub("_"," ",pheno$sp_ncbi)

```

## Stats

```{r}

# Gen
gen_sps <- unique(c(mt$spp, ms$spp, gen_d$spp, gen_lf$spp))

bsi_v1 <- intersect(biov1$spp, gen_sps)

# Break down
tot1 <- data.frame(N_species_with_gen_data = length(gen_sps),
                   v1_matches = length(bsi_v1))
tot1

# phenotypic
phen_sps <- unique(pheno$spp)

bsj_v1 <- intersect(biov1$spp, phen_sps)

# Break down
tot2 <- data.frame(N_species_with_pheno_data = length(phen_sps),
                   v1_matches = length(bsj_v1))
tot2

```

## Explore distribution

```{r}

# Malin
ggplot(mt, aes(He)) +
    geom_histogram()

ggplot(ms, aes(He)) +
    geom_histogram()

# De Kort
ggplot(gen_d, aes(GDp)) +
    geom_histogram()
# Why De Kort data has the two peaks? It is due to the marker type
ggplot(gen_d, aes(GDp)) +
    geom_histogram()+
    facet_grid(.~Marker)
# But if we look at the normalized GDp, the problem is solved. Therefore, we should be using the normalized version of GDp from DeKort
ggplot(gen_d, aes(GDp_norm)) +
    geom_histogram()

# Lawrence & Fraser
ggplot(gen_lf, aes(as.numeric(He))) +
    geom_histogram()

```

# Create a single genetic diversity dataset

```{r}

transform01 <- function(x) (x * (length(x) - 1) + 0.5) / (length(x))

# vec<-rnorm(100)+abs(min(vec))
# vec<-vec+min(vec)
# 
# hist(vec)
# hist(transform01(vec))

# Malin
mt_sub <- mt  %>%
    mutate(long = as.numeric(lon),
           lat = as.numeric(lat),
           N = as.numeric(n), # Sample size
           Marker = "Mitochondrial") %>% 
    select(spp, He, long, lat, N, Marker) %>%
    filter(!is.na(He) | !He==0)
mt_sub$Source = "Malin"
mt_sub$Het_type = "He"

ms_sub <- ms %>%
    mutate(long = as.numeric(lon),
           lat = as.numeric(lat),
           N = as.numeric(n), # Sample size
           Marker = "Microsatellite") %>% 
    select(spp, He, long, lat, N, Marker) %>%
    filter(!is.na(He) | !He==0)
ms_sub$Source = "Malin"
ms_sub$Het_type = "He"

###################
# De Kort

# Jonathan R.:
# co-dominant refers to the fact that you can distinguish homozigous and heterozygous individuals, such as microsatellites
# dominant means that you either have the detection of the marker or not. but you don’t know if the individual is heterozygous or not, such as AFLP
# restriction enzymes are proteins that you use to cut the genomes into fragments. there are used for AFLP and allozymes, so not sure

# De Kort code for Marker:
# CD = Co-dominant
# D = Dominant
# ENZ = Enzime

gen_d_sub <- gen_d %>%
    mutate(long = as.numeric(LONGITUDE),
           lat = as.numeric(LATITUDE),
           spp = spp_new, 
           He = as.numeric(GDp),
           N = as.numeric(SampleSize), # Sample size
           Marker = case_when(
               Marker=="CD" ~ "Microsatellite",
               Marker=="D" ~ "AFLP",
               Marker=="ENZ" ~ "AFLP",
               TRUE ~ as.character(Marker))) %>% 
    select(spp, He, long, lat, N, Marker) %>%
    filter(!is.na(He) | !He==0) %>%
    filter(!(Marker == "AFLP" & He > .5)) # AFLPs > 0.5 are errors
gen_d_sub$Source = "De Kort"
gen_d_sub$Het_type = "He"

# How DeKort calculated the normalized version of GDp
# gen_d_sub$He_harm <- NA
# 
# m <- unique(gen_d_sub$Marker)
# for(i in 1:length(m)){
#     pos <- which(gen_d_sub$Marker == m[i])
#     gen_d_sub$He_harm[pos] <-  scale(gen_d_sub$He[pos])
# }
# gen_d_sub$He_harm <-  (gen_d_sub$He_harm - min(gen_d_sub$He_harm))/(max(gen_d_sub$He_harm)-min(gen_d_sub$He_harm))

###################
# Lawrence & Fraser
gen_lf_sub <- gen_lf %>%
    mutate(long = as.numeric(Long),
           lat = as.numeric(Lat),
           He = as.numeric(He),
           Ho = as.numeric(Ho),
           spp = spp_new,
           N = as.numeric(n), # Sample size
           Marker = "Microsatellite") %>% 
    select(spp, He, Ho, long, lat, N, Marker) %>%
    filter((!is.na(He) & !is.na(Ho)) | !He==0 | !Ho==0)
gen_lf_sub$Source = "Lawrence & Fraser"
gen_lf_sub$Het_type = "He"
pos <- which(is.na(gen_lf_sub$He))
gen_lf_sub$He[pos] <- gen_lf_sub$Ho[pos]
gen_lf_sub$Het_type[pos] = "Ho"
gen_lf_sub <- gen_lf_sub %>% select(spp, He, Source, Het_type, long, lat, N, Marker)
gen_lf_sub <- gen_lf_sub %>%
    filter(He < 1)

###################
# Phenotypic rates
pheno_rate <- pheno %>%
    mutate(long = as.numeric(sample1.longitude),
           lat = as.numeric(sample1.latitude),
           trait1 = as.numeric(mean1),
           trait2 = as.numeric(mean2),
           spp = spp_new,
           Trait = trait_type, 
           Years = as.numeric(years),
           Gen = as.numeric(generations)) %>% 
    select(spp, long, lat, Trait, trait1, trait2, Years, Gen) %>%
    filter(!is.na(trait1) & !is.na(trait2))
pheno_rate$Source = "PROCEED"

## Calculate rate
for(i in 1:nrow(pheno_rate)){
    sppi <- pheno_rate$spp[i]
    traiti <- pheno_rate$Trait[i]
    pos_spp_trait_i <- which(pheno_rate$spp == sppi & pheno_rate$Trait == traiti)
    pheno_rate$PR1[i] <- (pheno_rate$trait2[i] - pheno_rate$trait1[i]) / mean(c(pheno_rate$trait2[pos_spp_trait_i], pheno_rate$trait1[pos_spp_trait_i])) * pheno_rate$Years[i]
    pheno_rate$PR2[i] <- (pheno_rate$trait2[i] - pheno_rate$trait1[i]) / sqrt(mean(c(pheno_rate$trait2[pos_spp_trait_i], pheno_rate$trait1[pos_spp_trait_i])))^2 * pheno_rate$Years[i]
}

###################
# Group all
gen_div <- rbind(mt_sub %>% select(names(mt_sub)), 
                 ms_sub %>% select(names(mt_sub)),
                 gen_d_sub %>% select(names(mt_sub)),
                 gen_lf_sub %>% select(names(mt_sub)))
gen_div <- na.omit(gen_div)

# View(gen_div %>%
#          group_by(spp) %>%
#          tally()) 

###################
# Harmonize following DeKort
# 1) scale values at each marker type
# 2) standardize (x - min)/(max-min)

gen_div$He_harm <- NA
gen_div$He_harm2 <- NA

m <- unique(gen_div$Marker)
for(i in 1:length(m)){
    pos <- which(gen_div$Marker == m[i])
    my_x <- scale(gen_div$He[pos])
    gen_div$He_harm[pos] <- (my_x - min(my_x, na.rm = T))/(max(my_x, na.rm = T)-min(my_x, na.rm = T))
    gen_div$He_harm2[pos] <-  transform01(gen_div$He[pos])
}

pheno_rate$PR_harm <- NA

m <- unique(pheno_rate$Trait)
for(i in 1:length(m)){
    pos <- which(pheno_rate$Trait == m[i])
    my_x <- scale(pheno_rate$PR1[pos])
    pheno_rate$PR_harm[pos] <-  (my_x - min(my_x, na.rm = T))/(max(my_x, na.rm = T)-min(my_x, na.rm = T))
}

# gen_div %>%
#     group_by(spp) %>%
#     tally() %>%
#     summary()

```

## Explore distribution of He

```{r}

# Full dataset
ggplot(gen_div, aes(He)) +
    geom_histogram()

ggplot(gen_div, aes(He_harm)) +
    geom_histogram()

ggplot(gen_div, aes(He_harm2)) +
    geom_histogram()

# Compare sources
ggplot(gen_div, aes(He)) +
    geom_histogram() +
    facet_wrap(.~Source, scales = "free_y")

ggplot(gen_div, aes(He_harm)) +
    geom_histogram() +
    facet_wrap(.~Source, scales = "free_y")

# Compare markers
ggplot(gen_div, aes(He)) +
    geom_histogram() +
    facet_wrap(.~Marker, scales = "free_y")

ggplot(gen_div, aes(He_harm)) +
    geom_histogram() +
    facet_wrap(.~Marker, scales = "free_y")

# Compare sources and markers
ggplot(gen_div, aes(He)) +
    geom_histogram() +
    facet_wrap(Marker~Source, scales = "free_y")

ggplot(gen_div, aes(He_harm)) +
    geom_histogram() +
    facet_wrap(Marker~Source, scales = "free_y")

# Look into phenotic rates
ggplot(pheno_rate, aes(PR1)) +
    geom_histogram() +
    facet_wrap(Trait~., scales = "free")

ggplot(pheno_rate, aes(PR_harm)) +
    geom_histogram() +
    facet_wrap(Trait~., scales = "free")

```

# Merge genetic and bioshifts data

```{r}

gen_data_v1 <- append(list(gen_div), 
                      list(biov1)) %>% reduce(full_join, by='spp')
gen_data_v1 <- gen_data_v1 %>% 
    filter(!is.na(He), !is.na(SHIFT))

# Remove groups with few species
rem <- gen_data_v1 %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))
rem <- rem$Class[which(rem$N<3)]

gen_data_v1 <- gen_data_v1 %>% 
    filter(!Class %in% rem)

# N species
length(unique(gen_data_v1$spp))

# Use only plus plus or minus minus
table(gen_data_v1$shift_vel_sign)

gen_data_v1 <- gen_data_v1 %>% filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

gen_data_v1
# N species
length(unique(gen_data_v1$spp))

# N obs per species
# View(table(gen_data_v1$spp))

ggplot(gen_data_v1, aes(He_harm2)) +
    geom_histogram()

ggplot(gen_data_v1, aes(He_harm)) +
    geom_histogram()

```

## Which species?

```{r}
toplot_ <- gen_data_v1 %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Class, y = N))+
    ggtitle("N species by Class")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r fig.width=10, fig.height=4}
gen_data_v1 %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1 %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r fig.width=10, fig.height=5}

gen_data_v1 %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1 %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

The merge we made considers all possible combinations of He and Shifts for each species. Therefore, there are lots of pseudo-replicates. Explore average difference in genetic diversity measures per species.

```{r}

DT::datatable(gen_div %>% 
                  group_by(spp) %>%
                  summarise(mean = mean(He_harm2),
                            sd = sd(He_harm2),
                            N = length(spp)))

# For some species we can get quite large variance in He. For example Fistularia commersonii:
gen_div %>% 
    filter(spp=="Fistularia commersonii")


```

# Mergind using average He per species

```{r}

gen_div_avg <- gen_div %>%
    group_by(spp) %>%
    summarise(N = length(spp),
              He = median(He_harm)) 

#merge all data frames in list
gen_data_v1_avg <- append(list(gen_div_avg), list(biov1)) %>% reduce(full_join, by='spp')
gen_data_v1_avg <- gen_data_v1_avg %>% filter(!is.na(He), !is.na(SHIFT))

# Remove groups with few species
rem <- gen_data_v1_avg %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))
rem <- rem$Class[which(rem$N<3)]

gen_data_v1_avg <- gen_data_v1_avg %>% 
    filter(!Class %in% rem)

# N species
length(unique(gen_data_v1_avg$spp))

# Use only plus plus or minus minus
table(gen_data_v1_avg$shift_vel_sign)

gen_data_v1_avg <- gen_data_v1_avg %>% filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

```

## Which species?

```{r}
toplot_ <- gen_data_v1_avg %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Class, y = N))+
    ggtitle("N species by Class")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r fig.width=10, fig.height=4}
gen_data_v1_avg %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_avg %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r fig.width=10, fig.height=5}

gen_data_v1_avg %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_avg %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

# Mergind using average He per Param & species

Use lower latitude TE and higher latitude as LE

```{r}

gen_div_avg_2 <- gen_div %>%
    group_by(spp) %>%
    summarise(#N = length(spp),
        LE = median(He_harm[which(lat==max(lat))]),
        O = median(He_harm),
        TE = median(He_harm[which(lat==min(lat))])) %>%
    # filter(N>5) %>%
    select(spp, LE, O, TE) %>%
    gather(Param, He, -spp)

#merge all data frames in list
gen_data_v1_avg_2 <- append(list(gen_div_avg_2), list(biov1)) %>% reduce(full_join, by=c("spp","Param"))
gen_data_v1_avg_2 <- gen_data_v1_avg_2 %>% filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_avg_2$spp))

# Remove groups with few species
rem <- gen_data_v1_avg_2 %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))
rem <- rem$Class[which(rem$N<3)]

gen_data_v1_avg_2 <- gen_data_v1_avg_2 %>% 
    filter(!Class %in% rem)

# N species
length(unique(gen_data_v1_avg_2$spp))

# Use only plus plus or minus minus
table(gen_data_v1_avg_2$shift_vel_sign)

# gen_data_v1_avg_2 <- gen_data_v1_avg_2 %>% filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_avg_2$spp))

```

## Which species?

```{r}
toplot_ <- gen_data_v1_avg_2 %>%
    group_by(Group) %>%
    summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r fig.width=10, fig.height=4}

gen_data_v1_avg_2 %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_avg_2 %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("Lag (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

```

## Per group

```{r fig.width=10, fig.height=5}

gen_data_v1_avg_2 %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_avg_2 %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

# Mergind based on distance

Get the avg gen div weighted by distance to the shift data

```{r echo=FALSE}

m <- unique(biov1$spp[which(biov1$spp %in% gen_div$spp)])

gen_data_v1_dist <- list()

for(i in 1:length(m)){ cat(i, "from", length(m), "\r")
    sp2go <- m[i]
    sub_gen <- subset(gen_div, spp == sp2go)
    sub_gen_v <- terra::vect(sub_gen, 
                             geom=c("long", "lat"), 
                             crs = "+proj=longlat +datum=WGS84 +no_defs")
    
    sub_bio <- subset(biov1, spp == sp2go)
    areas <- unique(sub_bio$ID)
    
    gen_data_v1_dist_tmp <- list()
    
    for(j in 1:length(areas)){
        area2go <- areas[j]
        sub_bio_area_j <- subset(sub_bio, ID == area2go)
        sub_bio_area_j_v <- terra::vect(unique(sub_bio_area_j[,c("long", "lat")]), 
                                        geom=c("long", "lat"), 
                                        crs = "+proj=longlat +datum=WGS84 +no_defs")
        
        # plot(terra::vect(rbind(sub_gen[,c("long", "lat")], sub_bio_area_j[,c("long", "lat")]),
        #                  geom=c("long", "lat"), 
        #                  crs = "+proj=longlat +datum=WGS84 +no_defs"))
        # plot(sub_bio_area_j_v, col = "red", add=T)
        
        dists <- terra::distance(sub_gen_v, sub_bio_area_j_v)[,1]
        
        # the closest
        dists_order <- order(dists)[1]
        # sub_gen <- sub_gen[dists_order,]
        
        # gen_data_v1_dist_tmp[[j]] <- append(list(sub_gen), list(sub_bio_area_j)) %>% 
        #     reduce(full_join, by="spp")
        
        # weighted distance
        wtd_he <- weighted.mean(sub_gen$He_harm, 1/dists) # inverse of the distance
        sub_gen_tmp <- sub_gen[dists_order,]
        sub_gen_tmp$He <- wtd_he
        
        gen_data_v1_dist_tmp[[j]] <- append(list(sub_gen_tmp), list(sub_bio_area_j)) %>% 
            reduce(full_join, by="spp")
    }
    
    gen_data_v1_dist[[i]] <- rbindlist(gen_data_v1_dist_tmp)
    
}

gen_data_v1_dist <- rbindlist(gen_data_v1_dist)

gen_data_v1_dist <- gen_data_v1_dist %>% filter(!is.na(He), !is.na(SHIFT))

# N species
length(unique(gen_data_v1_dist$spp))

# Remove groups with few species
rem <- gen_data_v1_dist %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))
rem <- rem$Class[which(rem$N<3)]

gen_data_v1_dist <- gen_data_v1_dist %>% 
    filter(!Class %in% rem)

# N species
length(unique(gen_data_v1_dist$spp))

# Use only plus plus or minus minus
table(gen_data_v1_dist$shift_vel_sign)

gen_data_v1_dist <- gen_data_v1_dist %>% filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

# N species
length(unique(gen_data_v1_dist$spp))

```

## Which species?

```{r}
toplot_ <- gen_data_v1_dist %>%
    group_by(Group) %>%
    summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Group, y = N))+
    ggtitle("N species by Group")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r fig.width=10, fig.height=4}
gen_data_v1_dist %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

gen_data_v1_dist %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r fig.width=10, fig.height=5}

gen_data_v1_dist %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

gen_data_v1_dist %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = He, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Genetic diversity (He)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```

# Playing with models

# Select a dataset

```{r}

mydataset <- gen_data_v1_dist

```

## 1 full models

### Latitude

#### Interaction in the fixed effect

Variables description:

lag3 = velocity-shift

lag3 = velocity-corrected_shift

lag33 = lag3 (= velocity when shifts in opposite direction)

lag34 = lag3 (= velocity when shifts in opposite direction)

```{r}

# What is the best model?
# lag3 model

lat_groups <- mydataset %>%
    filter(Type == "LAT") %>%
    group_by(Group) %>%
    summarise(N = length(unique(spp)))
lat_groups <- lat_groups$Group[which(lat_groups$N >= 5)]
lat_groups

mydatatogo <- mydataset %>%
    filter(Type == "LAT") %>%
    filter(Group %in% lat_groups) %>%
    mutate(He = scale(He),
           lag3 = scale(lag3),
           SHIFT = scale(SHIFT))

m_lat_l <- lmer(lag3 ~ He*Param + (1|ID) + (1|Group),
                REML=TRUE, na.action="na.fail",
                control = lmerControl(optimizer ='optimx', 
                                      optCtrl=list(method='nlminb')),
                data = mydatatogo)

MuMIn::r.squaredGLMM(m_lat_l)
# R2m       R2c
# 0.105442 0.6508075

summary(m_lat_l)

# Shift model
m_lat_s <- lmer(SHIFT ~ He*velocity*Param + (1|ID) + (1|Group),
                REML=TRUE, na.action="na.fail",
                control = lmerControl(optimizer ='optimx', 
                                      optCtrl=list(method='nlminb')),
                data = mydatatogo)

MuMIn::r.squaredGLMM(m_lat_s)
# R2m       R2c
# 0.3022617 0.6866355
summary(m_lat_s)

anova(m_lat_l,m_lat_s)
#          npar    AIC    BIC  logLik deviance  Chisq Df Pr(>Chisq)    
# m_lat_s   10 3939.4 3986.2 -1959.7   3919.4                         
# m_lat_l   22 3430.3 3533.1 -1693.1   3386.3 533.18 12  < 2.2e-16 ***
# The lag3 model is better

summary(m_lat_l)

plot_model(m_lat_l, type = "pred")
plot_model(m_lat_l, type = "pred", terms = c("He", "Param"))
plot_model(m_lat_l, type = "pred", terms = c("He", "Param", "Group"))


```

#### Interaction in the random effect

```{r}

# What is the best model?
# lag3 model
m_lat_l <- lmer(lag33 ~ He + (He|Param:Group) + (1|ID) + (1|Class),
                REML=TRUE,na.action="na.fail",
                control = lmerControl(optimizer ='optimx', 
                                      optCtrl=list(method='nlminb')),
                data = mydataset %>%
                    filter(Type == "LAT") %>%
                    filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_lat_l) 
# R2m       R2c
# 0.001001386 0.518533

# Shift model
m_lat_s <- lmer(SHIFT ~ He*velocity + (He|Param:Group) + (1|ID) + (1|Class),
                REML=TRUE,na.action="na.fail",
                control = lmerControl(optimizer ='optimx', 
                                      optCtrl=list(method='nlminb')),
                data = mydataset %>%
                    filter(Type == "LAT") %>%
                    filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_lat_s)
# R2m       R2c
# 0.01929231 0.3231063

anova(m_lat_l,m_lat_s)
# npar    AIC    BIC  logLik deviance  Chisq Df Pr(>Chisq)    
# m_lat_l    8 3448.2 3485.6 -1716.1   3432.2                    
# m_lat_s   10 3939.4 3986.2 -1959.7   3919.4     0  2          1
# No difference


summary(m_lat_s)

rr1 <- ranef(m_lat_s)
rr1dt <- as.data.frame(rr1)

ggplot(rr1dt %>%
           filter(grpvar == "Param:Group"), 
       aes(y=grp,x=condval)) +
    geom_point() + facet_wrap(~term,scales="free_x") +
    geom_errorbarh(aes(xmin=condval -2*condsd,
                       xmax=condval +2*condsd), height=0) +
    geom_vline(xintercept = 0, linetype = "dotted")

plot_model(m_lat_s, type = "pred", 
           pred.type = "re", # prediction is conditioned on random effects
           terms = c("He", "Param", "Group"))

```
#### Residual model
```{r}

# What is the best model?
# lag3 model
m_lat_l <- lm(lag34 ~ He*Param*Group, 
              data = mydataset %>%
                  filter(Type == "LAT") %>%
                  filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

tmp <- summary(m_lat_l)
tmp$r.squared
# R2 0.08499444

# Shift model
m_lat_s <- lm(shift_corrected ~ He*Param*Group*velocity,
              data = mydataset %>%
                  filter(Type == "LAT") %>%
                  filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

tmp <- summary(m_lat_s)
tmp$r.squared
# R2 0.09969969


AIC(m_lat_l) # 3610.265
AIC(m_lat_s) # 3872.998

# The shift model is better

summary(m_lat_s)

plot_model(m_lat_s, type = "pred", terms = c("He", "Param"))
plot_model(m_lat_s, type = "pred", terms = c("He", "Param", "Group"))

```

### Elevation

#### Interaction in the fixed effect

```{r}

# What is the best model?
# lag3 model
m_ele_l <- lmer(lag3 ~ He*Group*Param + (1|ID) + (1|spp),
                REML=TRUE,na.action="na.fail",
                control = lmerControl(optimizer ='optimx', 
                                      optCtrl=list(method='nlminb')),
                data = mydataset %>%
                    filter(Type == "ELE") %>%
                    filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_ele_l) 
# R2m       R2c
# 0.1639858 0.8498925

summary(m_ele_l)

newdata <- mydataset %>%
    filter(Type == "ELE") %>%
    filter(Group == "Bird" | Group == "Mammal" | Group == "Plant")

table(newdata$Group, newdata$Param)

# Shift model
m_ele_s <- lmer(SHIFT ~ He*Param*Group + velocity + DUR + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution +
                    (1|ID) + (1|Class/Family),
                REML=TRUE,na.action="na.fail",
                control=lmerControl(optimizer="Nelder_Mead",
                                    optCtrl=list(maxfun=1e4)),
                data = mydataset %>%
                    filter(Type == "ELE") %>%
                    filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_ele_s)
# R2m       R2c
# 0.1538334 0.7777449

anova(m_ele_l, m_ele_s)
# npar    AIC    BIC  logLik deviance  Chisq Df Pr(>Chisq)    
# m_ele_l   28 1924.9 2032.0 -934.46   1868.9                         
# m_ele_s   29 1875.4 1986.2 -908.68   1817.4 51.566  1  6.923e-13 ***

# The shift model is better

summary(m_ele_s)

plot_model(m_ele_l, type = "pred", terms = c("He", "Param"))
plot_model(m_ele_l, type = "pred", terms = c("He", "Param", "Group"))

```

#### Interaction in the random effect

```{r}

# What is the best model?
# lag3 model
m_ele_l <- lmer(lag3 ~ He + DUR + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution +
                    (He|Param:Group) + (1|ID) + (1|Class),
                REML=TRUE,na.action="na.fail",
                control=lmerControl(optimizer="Nelder_Mead",
                                    optCtrl=list(maxfun=1e4)),
                data = mydataset %>%
                    filter(Type == "ELE") %>%
                    filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_ele_l) 
# R2m       R2c
# 0.1404882 0.4756365

# Shift model
m_ele_s <- lmer(SHIFT ~ He + velocity + DUR + ID.area + Data + Sampling + Uncertainty_Parameter + Uncertainty_Distribution +
                    (He|Param:Group) + (1|ID) + (1|Class),
                REML=TRUE,na.action="na.fail",
                control=lmerControl(optimizer="Nelder_Mead",
                                    optCtrl=list(maxfun=1e4)),
                data = mydataset %>%
                    filter(Type == "ELE") %>%
                    filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

MuMIn::r.squaredGLMM(m_ele_s)
# R2m       R2c
# 0.04575894 0.2209294

anova(m_ele_l,m_ele_s)
# npar    AIC    BIC  logLik deviance  Chisq Df Pr(>Chisq)    
# m_ele_l   17 1926.1 1991.1 -946.05   1892.1                         
# m_ele_s   18 1878.4 1947.2 -921.21   1842.4 49.675  1  1.815e-12 ***

# The shift model is better

summary(m_ele_s)

rr1 <- ranef(m_ele_s)
rr1dt <- as.data.frame(rr1)

ggplot(rr1dt %>%
           filter(grpvar == "Param:Group"), 
       aes(y=grp,x=condval)) +
    geom_point() + facet_wrap(~term,scales="free_x") +
    geom_errorbarh(aes(xmin=condval -2*condsd,
                       xmax=condval +2*condsd), height=0) +
    geom_vline(xintercept = 0, linetype = "dotted")

plot_model(m_ele_s, type = "pred", 
           pred.type = "re", # prediction is conditioned on random effects
           terms = c("He", "Param", "Group"))
```

## 1 residual models

### Latitude

```{r}

# What is the best model?
# lag3 model
m_lat_l <- lm(lag3_corrected ~ He*Param*Group, 
              data = mydataset %>%
                  filter(Type == "LAT") %>%
                  filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

tmp <- summary(m_lat_l)
tmp$r.squared
# R2 0.06507956

# Shift model
m_lat_s <- lm(shift_corrected ~ He*Param*Group,
              data = mydataset %>%
                  filter(Type == "LAT") %>%
                  filter(Group == "Bird" | Group == "Fish" | Group == "Plant"))

tmp <- summary(m_lat_s)
tmp$r.squared
# R2 0.06244831


AIC(m_lat_l) # 3863.699
AIC(m_lat_s) # 3899.464

# The lag3 model is better

summary(m_lat_l)

plot_model(m_lat_l, type = "pred", terms = c("He", "Param"))
plot_model(m_lat_l, type = "pred", terms = c("He", "Param", "Group"))
```

### Elevation

```{r}

# What is the best model?
# lag3 model
m_ele_l <- lm(lag3_corrected ~ He*Param*Group, 
              data = mydataset %>%
                  filter(Type == "ELE") %>%
                  filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

tmp <- summary(m_ele_l)
tmp$r.squared
# R2 0.1576408

# Shift model
m_ele_s <- lm(shift_corrected ~ He*Param*Group,
              data = mydataset %>%
                  filter(Type == "ELE") %>%
                  filter(Group == "Bird" | Group == "Mammal" | Group == "Plant"))

tmp <- summary(m_ele_s)
tmp$r.squared
# R2 0.1704202


AIC(m_ele_l) # 1853.729
AIC(m_ele_s) # 1861.984

# The lag3 model is better

summary(m_ele_l)

plot_model(m_ele_l, type = "pred", terms = c("He", "Param"))
plot_model(m_ele_l, type = "pred", terms = c("He", "Param", "Group"))
```

# ----------------------------

# Phenotypic rates and bioshifts data

# Mergind using average He per species

```{r}

pheno_rate_avg <- pheno_rate %>%
    group_by(spp) %>%
    summarise(N = length(spp),
              PR = median(PR_harm)) 

#merge all data frames in list
pheno_rate_v1_avg <- append(list(pheno_rate_avg), list(biov1)) %>% reduce(full_join, by='spp')
pheno_rate_v1_avg <- pheno_rate_v1_avg %>% filter(!is.na(PR), !is.na(SHIFT))

# Remove groups with few species
rem <- pheno_rate_v1_avg %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))
rem <- rem$Class[which(rem$N<3)]

pheno_rate_v1_avg <- pheno_rate_v1_avg %>% 
    filter(!Class %in% rem)

# N species
length(unique(pheno_rate_v1_avg$spp))

# Use only plus plus or minus minus
table(pheno_rate_v1_avg$shift_vel_sign)

pheno_rate_v1_avg <- pheno_rate_v1_avg %>% filter(shift_vel_sign == "pospos" | shift_vel_sign == "negneg")

```

## Which species?

```{r}
toplot_ <- pheno_rate_v1_avg %>%
    group_by(Class) %>%
    summarise(N = length(unique(spp)))

ggplot(toplot_, aes(x = Class, y = N))+
    ggtitle("N species by Class")+
    geom_bar(stat="identity")+
    geom_text(aes(y=N, label=N), 
              hjust = -.1, vjust=0.2, size=3, 
              position = position_dodge(0.9))+
    theme_classic()+
    coord_flip()

```

## General pattern

```{r fig.width=10, fig.height=4}
pheno_rate_v1_avg %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = PR, y = lag3))+
    ggtitle("Latitude")+
    xlab("Phenotypic rates (PR)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)

pheno_rate_v1_avg %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = PR, y = lag3))+
    ggtitle("Elevation")+
    xlab("Phenotypic rates (PR)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_wrap(.~Param, scales = "free", ncol = 3)
```

## Per group

```{r fig.width=10, fig.height=5}

pheno_rate_v1_avg %>%
    filter(Type == "LAT") %>%
    ggplot(aes(x = PR, y = lag3, color = Group))+
    ggtitle("Latitude")+
    xlab("Phenotypic rates (PR)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")

pheno_rate_v1_avg %>%
    filter(Type == "ELE") %>%
    ggplot(aes(x = PR, y = lag3, color = Group))+
    ggtitle("Elevation")+
    xlab("Phenotypic rates (PR)")+
    ylab("lag3 (Velocity - Shift)") +
    geom_point(alpha = .5)+
    theme_bw()+
    geom_smooth(method = "lm")+
    facet_grid(Param~Group, scales = "free")


```
